### 2020-01-02 50sites.Rmd
added old code

---
title: "50sites"
author: "Yanlin Ma + Kevin Murgas"
date: "February 11, 2018"
output: word_document
---
Model 3: $$ y_{i,j,k} = \mu + b_j + I_{tumor} * ( \beta_T + c_j + d_k) + \epsilon_i $$
$$ b_j \sim N(0, \sigma_P), \ c_j \sim N(0,\sigma_{PT}),\ d_k \sim N(0, \sigma_T), \ \epsilon_i \sim N(0,\sigma_E)  $$
```{r echo=FALSE, include=FALSE}
library(rstan)
library(gtools)
library(ggplot2)
library(bayesplot)
```

```{r echo=FALSE, include=FALSE}
# Kevin's working directory
setwd("/Users/kevinmurgas/Documents/Data+ project/EPIC data")

# Yanlin's working directory
# setwd("D:/DataPlus2017/Data")

# load data
load("myFA.Rdata")

# Extract single site data
site <- function (site_no) {
  # extract CpG site xx to start
  temp <- FullAnnotation[site_no,]

  # here we use all patient samples, excluding glands
  indices <- c(9:13, 15:39,46,47,57,58,72:75)
  patientLabel <- substr(colnames(temp[indices]),1,1)
  patientLabel[10:12] <- "K*"
  sideLabel <- substr(colnames(temp[indices]),2,2)
  tissueLabel <- sideLabel
  tissueLabel[tissueLabel %in% c("A","B")] <- "T"   #replace A and B with T
  tumorIndicator <- 1*(tissueLabel=="T")

  work <- data.frame(beta=logit(t(temp[indices])), patient = patientLabel, tissue=tissueLabel, side=sideLabel, tInd=tumorIndicator)
  colnames(work)[1] <- "beta"
  return(work)
}

# Using Model PPTT (previously model 3): add intra-tumoral variance
stanfitPPTT <- function (dataset) {
  stanDat <- list(pID = as.integer(factor(dataset$patient)),
                  tInd = dataset$tInd,
                  N = nrow(dataset),
                  P = nlevels(dataset$patient),
                  y = dataset[,1])
  stanFitPPTT <-
    stan(
      #file = "model_PPTT.stan",
      fit = emptyFit,
      warmup = 1000,
      iter = 4000,
      chains = 4,
      thin = 1,
      data = stanDat,
      control = list(adapt_delta = 0.999),
      refresh = 0
    )
  return(stanFitPPTT=stanFitPPTT)
}

# Plot raw data and fits
fitplot <- function(data,stanFit) {
  # plot beta and fits at this site for all patients, using mean estimates for b and c coefficients
  p1 <- ggplot() + geom_point(data = data, aes(x=tInd, y=beta, colour=patient))
  pats <- as.integer(factor(data$patient))
  nsamp <- length(pats)
  npats <- max(pats)
  sumTemp <- summary(stanFit)$summary
  mu <- sumTemp[dim(sumTemp)[1]-5,1]
  betaT <- sumTemp[dim(sumTemp)[1]-6,1]
  sampcoefs <- data.frame(pat = pats, b = sumTemp[pats], c = sumTemp[pats+npats], d = sumTemp[1:nsamp+2*npats])
  patcoefs <- data.frame(pat = 1:npats, b = sumTemp[1:npats], c = sumTemp[1:npats+npats])
  rm(sumTemp)
  # add individual sample estimates
  sampdf <- data.frame(pat = rep(factor(data$patient),2), patg = rep(factor(1:nsamp),2), tInd = c(rep(0,nsamp),rep(1,nsamp)), est = c(sampcoefs$b+mu, (sampcoefs$b+sampcoefs$c++sampcoefs$d+mu+betaT)))
  p1 <- p1 + geom_line(data=sampdf, aes(x=tInd, y=est, colour=pat, group=patg), linetype="dashed", alpha=0.5)
  # add patient mean estimates
  patdf <- data.frame(pat = rep(levels(data$patient),2), tInd = c(rep(0,npats),rep(1,npats)), est = c(patcoefs$b+mu, patcoefs$b+patcoefs$c+mu+betaT))
  p1 <- p1 + geom_line(data=patdf, aes(x=tInd, y=est, colour=pat))
  # add overall mean estimates
  meandf <- data.frame(tInd = c(0,1), est = c(mu,mu+betaT))
  p1 <- p1 + geom_point(data=meandf, aes(x=tInd, y=est), size=3, alpha=0.7) + geom_line(data=meandf, aes(x=tInd, y=est), size=2, alpha=0.5)

  return(p1)
}

# Plot posterior distribution
postplot <- function(stanFit) {
  p2 <- mcmc_areas(
    as.array(stanFit), 
    pars = c("sigma_p", "sigma_pt","sigma_t","sigma_e"),
    prob = 0.8, # 80% intervals
    prob_outer = 0.95, 
    point_est = "median"
  )
  return(p2)
}

## CODE
# Randomly choose 50 sites
set.seed(50)
siteInds <- sample(1:866836,50)
nsites <- length(siteInds)

# check for model_PPTT.rds, must add to prevent crash
checkFile <- "model_PPTT.rds"
if (file.exists(checkFile)) {
  file.remove(checkFile)
  print("caught one!") 
}

# create empty fit to use same model compilation throughout
data1<-site(1)
stanDat1 <- list(
  pID = as.integer(factor(data1$patient)),
  tInd = data1$tInd,
  N = nrow(data1),
  P = nlevels(data1$patient),
  y = data1[, 1]
)
emptyFit <- stan(file="model_PPTT.stan", data = stanDat1, chains = 0)

```

```{r echo=FALSE,include=TRUE,error=FALSE,message=FALSE,warning=FALSE,results='hide',fig.keep='all',fig.show='hold',fig.height=3.7,fig.width=5}
# fig.align='center'
ptm <- proc.time()
for (i in siteInds) {
  #print(paste("Site ",i,sep=''))
  data <- site(i)
  stanFit <- stanfitPPTT(data)
  p1 <- fitplot(data, stanFit) + ggtitle(paste("Site", i, ": Beta Fits by Patient"))
  p2 <- postplot(stanFit) + ggtitle(paste("Site", i, ": Sigma Posteriors"))
  print(p1)
  print(p2)
}
mtp <- proc.time() - ptm

```

```{r echo=FALSE,include=TRUE}

print(mtp)
print(paste("Completed run on sites:", paste(siteInds, collapse=" ")))

```