### 2020-01-02 QC_and_Annotation.Rmd

---
title: "Quality Control"
output: word_document
---

## Density Plots of Beta Values

```{r, echo=FALSE, warning=FALSE,message=FALSE,fig.width=13,fig.height=5}
library(minfi)
library(IlluminaHumanMethylationEPICmanifest)
library(IlluminaHumanMethylationEPICanno.ilm10b2.hg19)
library(gdata) #gdata package needs Perl language installed
library(RColorBrewer)

# here set the working directory that points to the data folder
# e.g. the folder with all datasets in it, should contain all the
# 1337-1387 folders
# please just comment out the directories already here, so each
# user can uncomment the setwd for them as we move code around

# Kevin's working directory
# setwd("/Users/kevinmurgas/Documents/Data+ project/EPIC data")

# Yanlin's working directory
setwd("D:/DataPlus2017/Data")

####################### Function used for read in data #########################

# base_dir = directory containing basenames file
# targets = name of the dataset after reading in basenames file
# pat_dir = directory containing patient information file
# pat_file = WHOLE name (including directory) of the dataset after reading in patient information file
# work_name = name of the RGChannelSet

read.fun <- function(base_dir,pat_file) {
  # read in patient data from the .xlsx file
  pat_name <- read.xls(pat_file,header=TRUE)

  # Extract targets
  targets_name <- data.frame(pat_name[,"Complete.Barcode"])
  colnames(targets_name) <- "Basement"

  # Read in all .idat files
  targets_name$Basement <- file.path(base_dir, targets_name$Basement)
  work_name <- read.metharray(targets_name$Basement, verbose = TRUE)

  # Push the RGChannelSet to the global environment
  return(work_name)
}

############################### Read in data  #################################

## 1337 ##

base_dir_1337 <- "1337_Shibata EPIC DNA methylation data package/IDAT FILES"
pat_file_1337 <- "1337_Shibata EPIC DNA methylation data package/SAMPLE-ARRAY MAPPING/1337 (Shibata-8).xls"

work_1337 <- read.fun(base_dir_1337,pat_file_1337)

## 1345 ##

base_dir_1345 <- "1345_Shibata EPIC DNA methylation data package/IDAT FILES"
pat_file_1345 <- "1345_Shibata EPIC DNA methylation data package/SAMPLE-ARRAY MAPPING/1345 (Shibata-16).xlsx"

work_1345 <- read.fun(base_dir_1345,pat_file_1345)

# Two beadchips in plate 1345, read in seperately

pat_1345 <- read.xls(pat_file_1345)
targets_1345 <- data.frame(pat_1345[,"Complete.Barcode"])
colnames(targets_1345) <- 'Basement'

targets_1345$Basement <- file.path(base_dir_1345,targets_1345$Basement)
# Beadchip 1
work_1345a <- read.metharray(targets_1345$Basement[1:8],verbose=TRUE)
# Beadchip 2
work_1345b <- read.metharray(targets_1345$Basement[9:16],verbose = TRUE)

## 1350 ##

base_dir_1350 <- "1350_SHIBATA EPIC DNA METHYLATION DATA PACKAGE/IDAT files"
pat_file_1350 <- "1350_SHIBATA EPIC DNA METHYLATION DATA PACKAGE/SAMPLE-ARRAY MAPPING/1350 (Shibata-8).xlsx"

work_1350 <- read.fun(base_dir_1350,pat_file_1350)

## 1357 ##

base_dir_1357 <- "1357_Shibata EPIC DNA methylation data package/IDAT FILES"
pat_file_1357 <- "1357_Shibata EPIC DNA Methylation Data Package/SAMPLE-ARRAY MAPPING/1357 (Shibata-8).xlsx"

work_1357 <- read.fun(base_dir_1357,pat_file_1357)

## 1360 ##

base_dir_1360 <- "1360_Shibata EPIC Data Package/IDAT FILES"
pat_file_1360 <- "1360_Shibata EPIC Data Package/SAMPLE-ARRAY MAPPING/1360 (Shibata-8).xlsx"

work_1360 <- read.fun(base_dir_1360,pat_file_1360)

## 1378 ##

base_dir_1378 <- "1378_Shibata EPIC Data Package/IDAT FILES"
pat_file_1378 <- "1378_Shibata EPIC Data Package/SAMPLE-ARRAY MAPPING/1378 (Shbata-8).xls"

work_1378 <- read.fun(base_dir_1378,pat_file_1378)

## 1385 ##

base_dir_1385 <- "1385_Shibata EPIC Data Package/IDAT FILES"
pat_file_1385 <- "1385_Shibata EPIC Data Package/SAMPLE-ARRAY MAPPING/1385 (Shibata-8).xlsx"

work_1385 <- read.fun(base_dir_1385,pat_file_1385)

## 1387 ##

base_dir_1387 <- "1387_Shibata EPIC DNA methylation data package/IDAT FILES"
pat_file_1387 <- "1387_Shibata EPIC DNA methylation data package/SAMPLE-ARRAY MAPPING/1387 (Shibata-8).xls"

work_1387 <- read.fun(base_dir_1387,pat_file_1387)

############################# Bad Sample in 1337 ###############################

# Convert to MethylSet
mset_1337 <- preprocessRaw(work_1337)

# Check the methylation and unmethylation signals
# head(getMeth(mset_1337))
# head(getUnmeth((mset_1337)))

# Convert to RatioSet containing beta values and Mvalues(log(M/U))
rset_1337 <- ratioConvert(mset_1337,what = "both")

# Get beta
beta_1337 <- getBeta(rset_1337)

# Check beta values
# head(beta_1337)

#200360140022_R07C01 (JB) has bad quality
# dim(beta_1337)[1] == sum(is.na(beta_1337[,7]))

### read in new data without JB ###
base_dir_1337_new <- "1337_Shibata EPIC DNA methylation data package/IDAT FILES"

targets_1337_new <- read.csv("1337_Shibata EPIC DNA methylation data package/SAMPLE-ARRAY MAPPING/1337_Shibata-targets.csv", as.is = TRUE)

targets_1337_new$Basename <- file.path(base_dir_1337_new, targets_1337_new$Basename)
work_1337_new <- read.metharray(targets_1337_new$Basename, verbose = TRUE)

######################## Preprocessing/Normalization ##########################

### 1337 ###

## noob ##
noob_1337 <- preprocessNoob(work_1337_new)

rset_1337_noob <- ratioConvert(noob_1337,what="both")
beta_1337_noob <- getBeta(rset_1337_noob)

# check
# head(beta_1337_noob)
# Compare with raw beta values
# head(beta_1337)

## Genome Studio ##
lumi_1337 <- preprocessIllumina(work_1337_new,bg.correct = TRUE,
                                normalize = "controls",reference = 2)
rset_1337_lumi <- ratioConvert(lumi_1337,what="both")
beta_1337_lumi <- getBeta(rset_1337_lumi)

# check
# head(beta_1337_lumi)

# Compare with noob and raw
# head(beta_1337)
# head(beta_1337_noob)

### 1345 ###

## noob ##
noob_1345 <- preprocessNoob(work_1345)
noob_1345a <- preprocessNoob(work_1345a)
noob_1345b <- preprocessNoob(work_1345b)

## Genome Studio ##
lumi_1345 <- preprocessIllumina(work_1345,bg.correct = TRUE,
                                normalize = "controls",reference = 2)
lumi_1345a <- preprocessIllumina(work_1345a,bg.correct = TRUE,
                                normalize = "controls",reference = 2)
lumi_1345b <- preprocessIllumina(work_1345b,bg.correct = TRUE,
                                normalize = "controls",reference = 2)

### 1350 ###

## noob ##
noob_1350 <- preprocessNoob(work_1350)

## Genome Studio ##
lumi_1350 <- preprocessIllumina(work_1350,bg.correct = TRUE,
                                normalize = "controls",reference = 2)

### 1357 ###

## noob ##
noob_1357 <- preprocessNoob(work_1357)

## Genome Studio ##
lumi_1357 <- preprocessIllumina(work_1357,bg.correct = TRUE,
                                normalize = "controls",reference = 2)

### 1360 ###

## noob ##
noob_1360 <- preprocessNoob(work_1360)

## Genome Studio ##
lumi_1360 <- preprocessIllumina(work_1360,bg.correct = TRUE,
                                normalize = "controls",reference = 2)

### 1378 ###

## noob ##
noob_1378 <- preprocessNoob(work_1378)

## Genome Studio ##
lumi_1378 <- preprocessIllumina(work_1378,bg.correct = TRUE,
                                normalize = "controls",reference = 2)

### 1385 ###

## noob ##
noob_1385 <- preprocessNoob(work_1385)

## Genome Studio ##
lumi_1385 <- preprocessIllumina(work_1385,bg.correct = TRUE,
                                normalize = "controls",reference = 2)

### 1387 ###

## noob ##
noob_1387 <- preprocessNoob(work_1387)

## Genome Studio ##
lumi_1387 <- preprocessIllumina(work_1387,bg.correct = TRUE,
                                normalize = "controls",reference = 2)

############################### Quality Control ################################

####### Density Plots #########

## 1337 ##

pat_1337 <- read.xls("1337_Shibata EPIC DNA methylation data package/SAMPLE-ARRAY MAPPING/1337 (Shibata-8).xls")

par(mfrow=c(1,2),mar=c(4,2,4,1))
densityPlot(noob_1337,sampGroups = pat_1337$Sample_No,
            main = "Corrected Beta Values Using 'noob'",legend = FALSE)
legend("topright",legend = levels(pat_1337$Sample_No),
       text.col = brewer.pal(8,"Dark2"),cex=0.5)
densityPlot(lumi_1337,sampGroups = pat_1337$Sample_No,
            main = "Corrected Beta Values Using 'Illumina'",legend = FALSE)
legend("topright",legend = levels(pat_1337$Sample_No),
       text.col = brewer.pal(8,"Dark2"),cex=0.5)

## 1345 ##
pat_1345 <- read.xls("1345_Shibata EPIC DNA methylation data package/SAMPLE-ARRAY MAPPING/1345 (Shibata-16).xlsx")

par(mfrow=c(1,2),mar=c(4,2,4,1))
densityPlot(noob_1345,sampGroups = pat_1345$SAMPLE.ID,
            main = "Corrected Beta Values Using 'noob'",legend = FALSE)
legend("top",legend = levels(pat_1345$SAMPLE.ID),
       text.col = brewer.pal(8,"Dark2"),cex=0.5)
densityPlot(lumi_1345,sampGroups = pat_1345$SAMPLE.ID,
            main = "Corrected Beta Values Using 'Illumina'",legend = FALSE)
legend("top",legend = levels(pat_1345$SAMPLE.ID),
       text.col = brewer.pal(8,"Dark2"),cex=0.5)

## 1350 ##
pat_1350 <- read.xls("1350_SHIBATA EPIC DNA METHYLATION DATA PACKAGE/SAMPLE-ARRAY MAPPING/1350 (Shibata-8).xlsx")

par(mfrow=c(1,2),mar=c(4,2,4,1))
densityPlot(noob_1350,sampGroups = pat_1350$SAMPLE.ID,
            main = "Corrected Beta Values Using 'noob'",legend = FALSE)
legend("topright",legend = levels(pat_1350$SAMPLE.ID),
       text.col = brewer.pal(8,"Dark2"),cex=0.7)
densityPlot(lumi_1350,sampGroups = pat_1350$SAMPLE.ID,
            main = "Corrected Beta Values Using 'Illumina'",legend = FALSE)
legend("topright",legend = levels(pat_1350$SAMPLE.ID),
       text.col = brewer.pal(8,"Dark2"),cex=0.7)

## 1357 ## 
pat_1357 <- read.xls("1357_Shibata EPIC DNA Methylation Data Package/SAMPLE-ARRAY MAPPING/1357 (Shibata-8).xlsx")

par(mfrow=c(1,2),mar=c(4,2,4,1))
densityPlot(noob_1357,sampGroups = pat_1357$Sample.ID,
            main = "Corrected Beta Values Using 'noob'",legend = FALSE)
legend("top",legend = levels(pat_1357$Sample.ID),
       text.col = brewer.pal(8,"Dark2"),cex=0.7)
densityPlot(lumi_1357,sampGroups = pat_1357$Sample.ID,
            main = "Corrected Beta Values Using 'Illumina'",legend = FALSE)
legend("top",legend = levels(pat_1357$Sample.ID),
       text.col = brewer.pal(8,"Dark2"),cex=0.7)

## 1360 ##
pat_1360 <- read.xls("1360_Shibata EPIC Data Package/SAMPLE-ARRAY MAPPING/1360 (Shibata-8).xlsx")

par(mfrow=c(1,2),mar=c(4,2,4,1))
densityPlot(noob_1360,sampGroups = pat_1360$Sample.ID,
            main = "Corrected Beta Values Using 'noob'",legend = FALSE)
legend("top",legend = levels(pat_1360$Sample.ID),
       text.col = brewer.pal(8,"Dark2"),cex=0.7)
densityPlot(lumi_1360,sampGroups = pat_1360$Sample.ID,
            main = "Corrected Beta Values Using 'Illumina'",legend = FALSE)
legend("top",legend = levels(pat_1360$Sample.ID),
       text.col = brewer.pal(8,"Dark2"),cex=0.7)

## 1378 ##

pat_1378 <- read.xls("1378_Shibata EPIC Data Package/SAMPLE-ARRAY MAPPING/1378 (Shbata-8).xls")

par(mfrow=c(1,2),mar=c(4,2,4,1))
densityPlot(noob_1378,sampGroups = pat_1378$Tube.Label,
            main = "Corrected Beta Values Using 'noob'",legend = FALSE)
legend("top",legend = levels(pat_1378$Tube.Label),
       text.col = brewer.pal(8,"Dark2"),cex=0.7)
densityPlot(lumi_1378,sampGroups = pat_1378$Tube.Label,
            main = "Corrected Beta Values Using 'Illumina'",legend = FALSE)
legend("top",legend = levels(pat_1378$Tube.Label),
       text.col = brewer.pal(8,"Dark2"),cex=0.7)

## 1385 ##

pat_1385 <- read.xls("1385_Shibata EPIC Data Package/SAMPLE-ARRAY MAPPING/1385 (Shibata-8).xlsx")

par(mfrow=c(1,2),mar=c(4,2,4,1))
densityPlot(noob_1385,sampGroups = pat_1385$Tube.Label,
            main = "Corrected Beta Values Using 'noob'",legend = FALSE)
legend("top",legend = levels(pat_1385$Tube.Label),
       text.col = brewer.pal(8,"Dark2"),cex=0.7)
densityPlot(lumi_1385,sampGroups = pat_1385$Tube.Label,
            main = "Corrected Beta Values Using 'Illumina'",legend = FALSE)
legend("top",legend = levels(pat_1385$Tube.Label),
       text.col = brewer.pal(8,"Dark2"),cex=0.7)

## 1387 ##

pat_1387 <- read.xls("1387_Shibata EPIC DNA methylation data package/SAMPLE-ARRAY MAPPING/1387 (Shibata-8).xls")

par(mfrow=c(1,2),mar=c(4,2,4,1))
densityPlot(noob_1387,sampGroups = pat_1387$Sample.ID,
            main = "Corrected Beta Values Using 'noob'",legend = FALSE)
legend("top",legend = levels(pat_1387$Sample.ID),
       text.col = brewer.pal(8,"Dark2"),cex=0.7)
densityPlot(lumi_1387,sampGroups = pat_1387$Sample.ID,
            main = "Corrected Beta Values Using 'Illumina'",legend = FALSE)
legend("top",legend = levels(pat_1387$Sample.ID),
       text.col = brewer.pal(8,"Dark2"),cex=0.7)

```

``` {r,echo=FALSE, warning=FALSE,message=FALSE,eval=FALSE}

######### Median Plot (noob) ############

qc_1337_noob <- getQC(noob_1337)
qc_1345_noob <- getQC(noob_1345)
qc_1350_noob <- getQC(noob_1350)
qc_1357_noob <- getQC(noob_1357)
qc_1360_noob <- getQC(noob_1360)
qc_1378_noob <- getQC(noob_1378)
qc_1385_noob <- getQC(noob_1385)
qc_1387_noob <- getQC(noob_1387)

qc_noob <- rbind(qc_1337_noob,qc_1345_noob,qc_1350_noob,qc_1357_noob,qc_1360_noob, qc_1378_noob,qc_1385_noob,qc_1387_noob)

par(mfrow=c(1,1),mar=c(5,5,5,5))
plotQC(qc_noob,badSampleCutoff = 9)

######### Median Plot (Illumina) ############

qc_1337_lumi <- getQC(lumi_1337)
qc_1345_lumi <- getQC(lumi_1345)
qc_1350_lumi <- getQC(lumi_1350)
qc_1357_lumi <- getQC(lumi_1357)
qc_1360_lumi <- getQC(lumi_1360)
qc_1378_lumi <- getQC(lumi_1378)
qc_1385_lumi <- getQC(lumi_1385)
qc_1387_lumi <- getQC(lumi_1387)

qc_lumi <- rbind(qc_1337_lumi,qc_1345_lumi,qc_1350_lumi,qc_1357_lumi,qc_1360_lumi,
                 qc_1378_lumi,qc_1385_lumi,qc_1387_lumi)

par(mfrow=c(1,1),mar=c(5,5,5,5))
plotQC(qc_lumi,badSampleCutoff = 9)

```

## Multidimensional Scaling/Principle Component Analysis Plot

``` {r,echo=FALSE, eval=FALSE,warning=FALSE,message=FALSE,fig.width=17,fig.height=10}

############################### With JA ########################################

# Make sure var names are consistent
colnames(pat_1337)[8] <- "Sample.ID"
colnames(pat_1345)[8] <- "Sample.ID"
colnames(pat_1350)[8] <- "Sample.ID"
colnames(pat_1357)[8] <- "Sample.ID"
colnames(pat_1360)[8] <- "Sample.ID"
colnames(pat_1378)[8] <- "Sample.ID"
colnames(pat_1385)[8] <- "Sample.ID"
colnames(pat_1387)[8] <- "Sample.ID"

# Create a data frame that contains all patient information
pat <- rbind(pat_1337[,1:8],pat_1345[,1:8],pat_1350[,1:8],pat_1357[,1:8],
             pat_1360[,1:8],pat_1378[,1:8],pat_1385[,1:8],pat_1387[,1:8])

# Create a var indicating adenomas/carcinomas in pat
for(i in (1:dim(pat)[1])) {

     if ((pat$Sample.ID[i] == "KA") | (pat$Sample.ID[i] == "KB") |
      pat$Sample.ID[i] == "XA" | pat$Sample.ID[i] == "XB" |
      pat$Sample.ID[i] == "SA" | pat$Sample.ID[i] == "SB" |
      pat$Sample.ID[i] == "PA" | pat$Sample.ID[i] == "PB" |
      pat$Sample.ID[i] == "KA LL" | pat$Sample.ID[i] == "KB LL") {

       pat$type[i] <- "Adenomas"

      } else if (pat$Sample.ID[i] == "HA" | pat$Sample.ID[i] == "HB" |
                 pat$Sample.ID[i] == "HN" | pat$Sample.ID[i] == "JA" |
                 pat$Sample.ID[i] == "JN" | pat$Sample.ID[i] == "KA" |
                 pat$Sample.ID[i] == "EA" | pat$Sample.ID[i] == "EB" |
                 pat$Sample.ID[i] == "EN" | pat$Sample.ID[i] == "KA" |
                 pat$Sample.ID[i] == "KA (new)" | 
                 pat$Sample.ID[i] == "KB (new)" | 
                 pat$Sample.ID[i] == "KN (new)" | 
                 pat$Sample.ID[i] == "WA" | pat$Sample.ID[i] == "WB" |
                 pat$Sample.ID[i] == "WN" | pat$Sample.ID[i] == "TA" |
                 pat$Sample.ID[i] == "TB" | pat$Sample.ID[i] == "JB" |
                 pat$Sample.ID[i] == "KA" | pat$Sample.ID[i] == "CA" |
                 pat$Sample.ID[i] == "CB" | pat$Sample.ID[i] == "CN" |
                 pat$Sample.ID[i] == "MA" | pat$Sample.ID[i] == "MB" |
                 pat$Sample.ID[i] == "JA" | pat$Sample.ID[i] == "OA" |
                 pat$Sample.ID[i] == "OB" | pat$Sample.ID[i] == "FA" |
                 pat$Sample.ID[i] == "FB" | pat$Sample.ID[i] == "KA" |
                 pat$Sample.ID[i] == "CA LL" | pat$Sample.ID[i] == "CB LL" |
                 pat$Sample.ID[i] == "FA LL" | pat$Sample.ID[i] == "FB LL" |
                 pat$Sample.ID[i] == "DA" | pat$Sample.ID[i] == "DB" |
                 pat$Sample.ID[i] == "UA" | pat$Sample.ID[i] == "UB" |
                 pat$Sample.ID[i] == "HALL-1" | pat$Sample.ID[i] == "HALL-2"|
                 pat$Sample.ID[i] == "HBLL-3" | pat$Sample.ID[i] == "HBLL-4") {

         pat$type[i] <- "Carcinomas"

      } else {
         pat$type[i] <- "Other"
        }
}

## PCA plot ##
memory.limit(size=60000)
all_noob <- combineArrays(noob_1337,noob_1345)
all_noob <- combineArrays(all_noob,noob_1350)
all_noob <- combineArrays(all_noob,noob_1357)
all_noob <- combineArrays(all_noob,noob_1360)
all_noob <- combineArrays(all_noob,noob_1378)
all_noob <- combineArrays(all_noob,noob_1385)
all_noob <- combineArrays(all_noob,noob_1387)

all_lumi <- combineArrays(lumi_1337,lumi_1345)
all_lumi <- combineArrays(all_lumi,lumi_1350)
all_lumi <- combineArrays(all_lumi,lumi_1357)
all_lumi <- combineArrays(all_lumi,lumi_1360)
all_lumi <- combineArrays(all_lumi,lumi_1378)
all_lumi <- combineArrays(all_lumi,lumi_1385)
all_lumi <- combineArrays(all_lumi,lumi_1387)

par(mfrow=c(1,2),mar=c(5,3,5,3))
mdsPlot(all_noob,numPositions = 1000000,sampGroups = pat$type,
        main = "PCA Plot Using Noob (including JA)", pch=16)

mdsPlot(all_lumi,numPositions = 1000000,sampGroups = pat$type,
        main = "PCA Plot Using Illumina (including JA)", pch = 16)

```

``` {r,echo=FALSE,warning=FALSE,message=FALSE,fig.width=17,fig.height=10}
##################################### Without JA ########################################

### read in new data without JA ###
base_dir_1337_new2 <- "1337_Shibata EPIC DNA methylation data package/IDAT FILES"

targets_1337_new2 <- read.csv("1337_Shibata EPIC DNA methylation data package/SAMPLE-ARRAY MAPPING/1337_Shibata-targets.csv", as.is = TRUE)

targets_1337_new2 <- targets_1337_new2[c(1:5,7),]
targets_1337_new2 <- file.path(base_dir_1337_new, targets_1337_new2)
work_1337_new2 <- read.metharray(targets_1337_new2, verbose = TRUE)

# normalization

## noob ##
noob_1337_new <- preprocessNoob(work_1337_new2)

## Genome Studio ##
lumi_1337_new <- preprocessIllumina(work_1337_new2,bg.correct = TRUE,
                                normalize = "controls",reference = 2)

# Make sure var names are consistent
colnames(pat_1337)[8] <- "Sample.ID"
colnames(pat_1345)[8] <- "Sample.ID"
colnames(pat_1350)[8] <- "Sample.ID"
colnames(pat_1357)[8] <- "Sample.ID"
colnames(pat_1360)[8] <- "Sample.ID"
colnames(pat_1378)[8] <- "Sample.ID"
colnames(pat_1385)[8] <- "Sample.ID"
colnames(pat_1387)[8] <- "Sample.ID"

# Create a data frame that contains all patient information
pat <- rbind(pat_1337[,1:8],pat_1345[,1:8],pat_1350[,1:8],pat_1357[,1:8],
             pat_1360[,1:8],pat_1378[,1:8],pat_1385[,1:8],pat_1387[,1:8])

# Create a var indicating adenomas/carcinomas in pat
for(i in (1:dim(pat)[1])) {

     if ((pat$Sample.ID[i] == "KA") | (pat$Sample.ID[i] == "KB") |
      pat$Sample.ID[i] == "XA" | pat$Sample.ID[i] == "XB" |
      pat$Sample.ID[i] == "SA" | pat$Sample.ID[i] == "SB" |
      pat$Sample.ID[i] == "PA" | pat$Sample.ID[i] == "PB" |
      pat$Sample.ID[i] == "KA LL" | pat$Sample.ID[i] == "KB LL") {

       pat$type[i] <- "Adenomas"

      } else if (pat$Sample.ID[i] == "HA" | pat$Sample.ID[i] == "HB" |
                 pat$Sample.ID[i] == "HN" | pat$Sample.ID[i] == "JA" |
                 pat$Sample.ID[i] == "JN" | pat$Sample.ID[i] == "KA" |
                 pat$Sample.ID[i] == "EA" | pat$Sample.ID[i] == "EB" |
                 pat$Sample.ID[i] == "EN" | pat$Sample.ID[i] == "KA" |
                 pat$Sample.ID[i] == "KA (new)" | 
                 pat$Sample.ID[i] == "KB (new)" | 
                 pat$Sample.ID[i] == "KN (new)" | 
                 pat$Sample.ID[i] == "WA" | pat$Sample.ID[i] == "WB" |
                 pat$Sample.ID[i] == "WN" | pat$Sample.ID[i] == "TA" |
                 pat$Sample.ID[i] == "TB" | pat$Sample.ID[i] == "JB" |
                 pat$Sample.ID[i] == "KA" | pat$Sample.ID[i] == "CA" |
                 pat$Sample.ID[i] == "CB" | pat$Sample.ID[i] == "CN" |
                 pat$Sample.ID[i] == "MA" | pat$Sample.ID[i] == "MB" |
                 pat$Sample.ID[i] == "JA" | pat$Sample.ID[i] == "OA" |
                 pat$Sample.ID[i] == "OB" | pat$Sample.ID[i] == "FA" |
                 pat$Sample.ID[i] == "FB" | pat$Sample.ID[i] == "KA" |
                 pat$Sample.ID[i] == "CA LL" | pat$Sample.ID[i] == "CB LL" |
                 pat$Sample.ID[i] == "FA LL" | pat$Sample.ID[i] == "FB LL" |
                 pat$Sample.ID[i] == "DA" | pat$Sample.ID[i] == "DB" |
                 pat$Sample.ID[i] == "UA" | pat$Sample.ID[i] == "UB" |
                 pat$Sample.ID[i] == "HALL-1" | pat$Sample.ID[i] == "HALL-2"|
                 pat$Sample.ID[i] == "HBLL-3" | pat$Sample.ID[i] == "HBLL-4") {

         pat$type[i] <- "Carcinomas"

      } else {
         pat$type[i] <- "Other"
        }
}

## PCA plot ##
memory.limit(size=30000)
all_noob <- combineArrays(noob_1337_new,noob_1345)
all_noob <- combineArrays(all_noob,noob_1350)
all_noob <- combineArrays(all_noob,noob_1357)
all_noob <- combineArrays(all_noob,noob_1360)
all_noob <- combineArrays(all_noob,noob_1378)
all_noob <- combineArrays(all_noob,noob_1385)
all_noob <- combineArrays(all_noob,noob_1387)

all_lumi <- combineArrays(lumi_1337_new,lumi_1345)
all_lumi <- combineArrays(all_lumi,lumi_1350)
all_lumi <- combineArrays(all_lumi,lumi_1357)
all_lumi <- combineArrays(all_lumi,lumi_1360)
all_lumi <- combineArrays(all_lumi,lumi_1378)
all_lumi <- combineArrays(all_lumi,lumi_1385)
all_lumi <- combineArrays(all_lumi,lumi_1387)

par(mfrow=c(1,2),mar=c(5,3,5,3))
mdsPlot(all_noob,numPositions = 1000000,sampGroups = pat$type,
        main = "PCA Plot Using Noob (without JA)", pch=16)

mdsPlot(all_lumi,numPositions = 1000000,sampGroups = pat$type,
        main = "PCA Plot Using Illumina (without JA)", pch=16)

```

``` {r, echo=FALSE,warning=FALSE,message=FALSE,fig.width=17,fig.height=10}
####################################### Without Cell Lines and Xenografts #############################

####### 1360: all 
####### 1378: two

### read in 1378 without cell lines and xenografts###
base_dir_1378_new <- "1378_Shibata EPIC Data Package/IDAT FILES"

targets_1378_new <- read.xls("1378_Shibata EPIC Data Package/SAMPLE-ARRAY MAPPING/1378 (Shbata-8).xls")

targets_1378_new <- targets_1378_new[c(-1,-4),7]
targets_1378_new <- file.path(base_dir_1378_new, targets_1378_new)
work_1378_new <- read.metharray(targets_1378_new, verbose = TRUE)

# normalization

## noob ##
noob_1378_new <- preprocessNoob(work_1378_new)

## Genome Studio ##
lumi_1378_new <- preprocessIllumina(work_1378_new,bg.correct = TRUE,
                                normalize = "controls",reference = 2)

# Make sure var names are consistent
colnames(pat_1337)[8] <- "Sample.ID"
colnames(pat_1345)[8] <- "Sample.ID"
colnames(pat_1350)[8] <- "Sample.ID"
colnames(pat_1357)[8] <- "Sample.ID"
colnames(pat_1360)[8] <- "Sample.ID"
colnames(pat_1378)[8] <- "Sample.ID"
colnames(pat_1385)[8] <- "Sample.ID"
colnames(pat_1387)[8] <- "Sample.ID"

# Create a data frame that contains all patient information
pat <- rbind(pat_1337[,1:8],pat_1345[,1:8],pat_1350[,1:8],pat_1357[,1:8],
             pat_1360[,1:8],pat_1378[,1:8],pat_1385[,1:8],pat_1387[,1:8])

pat <- pat[-c(33:38,41:49,52),]

# Create a var indicating adenomas/carcinomas in pat
for(i in (1:dim(pat)[1])) {

     if ((pat$Sample.ID[i] == "KA") | (pat$Sample.ID[i] == "KB") |
      pat$Sample.ID[i] == "XA" | pat$Sample.ID[i] == "XB" |
      pat$Sample.ID[i] == "SA" | pat$Sample.ID[i] == "SB" |
      pat$Sample.ID[i] == "PA" | pat$Sample.ID[i] == "PB" |
      pat$Sample.ID[i] == "KA LL" | pat$Sample.ID[i] == "KB LL") {

       pat$type[i] <- "Adenomas"

      } else if (pat$Sample.ID[i] == "HA" | pat$Sample.ID[i] == "HB" |
                 pat$Sample.ID[i] == "HN" | pat$Sample.ID[i] == "JA" |
                 pat$Sample.ID[i] == "JN" | pat$Sample.ID[i] == "KA" |
                 pat$Sample.ID[i] == "EA" | pat$Sample.ID[i] == "EB" |
                 pat$Sample.ID[i] == "EN" | pat$Sample.ID[i] == "KA" |
                 pat$Sample.ID[i] == "KA (new)" | 
                 pat$Sample.ID[i] == "KB (new)" | 
                 pat$Sample.ID[i] == "KN (new)" | 
                 pat$Sample.ID[i] == "WA" | pat$Sample.ID[i] == "WB" |
                 pat$Sample.ID[i] == "WN" | pat$Sample.ID[i] == "TA" |
                 pat$Sample.ID[i] == "TB" | pat$Sample.ID[i] == "JB" |
                 pat$Sample.ID[i] == "KA" | pat$Sample.ID[i] == "CA" |
                 pat$Sample.ID[i] == "CB" | pat$Sample.ID[i] == "CN" |
                 pat$Sample.ID[i] == "MA" | pat$Sample.ID[i] == "MB" |
                 pat$Sample.ID[i] == "JA" | pat$Sample.ID[i] == "OA" |
                 pat$Sample.ID[i] == "OB" | pat$Sample.ID[i] == "FA" |
                 pat$Sample.ID[i] == "FB" | pat$Sample.ID[i] == "KA" |
                 pat$Sample.ID[i] == "CA LL" | pat$Sample.ID[i] == "CB LL" |
                 pat$Sample.ID[i] == "FA LL" | pat$Sample.ID[i] == "FB LL" |
                 pat$Sample.ID[i] == "DA" | pat$Sample.ID[i] == "DB" |
                 pat$Sample.ID[i] == "UA" | pat$Sample.ID[i] == "UB" |
                 pat$Sample.ID[i] == "HALL-1" | pat$Sample.ID[i] == "HALL-2"|
                 pat$Sample.ID[i] == "HBLL-3" | pat$Sample.ID[i] == "HBLL-4") {

         pat$type[i] <- "Carcinomas"

      } 
}

## PCA plot ##
memory.limit(size=30000)
all_noob <- combineArrays(noob_1337_new,noob_1345)
all_noob <- combineArrays(all_noob,noob_1350)
all_noob <- combineArrays(all_noob,noob_1357)
all_noob <- combineArrays(all_noob,noob_1378_new)
all_noob <- combineArrays(all_noob,noob_1385)
all_noob <- combineArrays(all_noob,noob_1387)

all_lumi <- combineArrays(lumi_1337_new,lumi_1345)
all_lumi <- combineArrays(all_lumi,lumi_1350)
all_lumi <- combineArrays(all_lumi,lumi_1357)
all_lumi <- combineArrays(all_lumi,lumi_1378_new)
all_lumi <- combineArrays(all_lumi,lumi_1385)
all_lumi <- combineArrays(all_lumi,lumi_1387)

par(mfrow=c(1,2),mar=c(5,3,5,3))
mdsPlot(all_noob,numPositions = 1000000,sampGroups = pat$type,
        main = "PCA Plot Using Noob (without JA)", pch=16)

mdsPlot(all_lumi,numPositions = 1000000,sampGroups = pat$type,
        main = "PCA Plot Using Illumina (without JA)", pch=16)

```

## Predicting Sex

``` {r,echo=FALSE, warning=FALSE,message=FALSE}

all_gmset <- mapToGenome(all_noob)
all_sex <- getSex(all_gmset)
plotSex(all_sex)

```

## Percent of Probes

```{r,echo=FALSE, warning=FALSE}

########################### Percent of probes ################################

per_fun <- function (work) {

  mset <- preprocessRaw(work)

  rset <- ratioConvert(mset,what = "both")

  beta <- getBeta(rset)

  per <- 1-apply(is.na(beta),2,sum)/dim(beta)[1]

  return(per)
}

per_1337 <- per_fun(work_1337)
per_1345 <- per_fun(work_1345)
per_1350 <- per_fun(work_1350)
per_1357 <- per_fun(work_1357)
per_1360 <- per_fun(work_1360)
per_1378 <- per_fun(work_1378)
per_1385 <- per_fun(work_1385)
per_1387 <- per_fun(work_1387)

```

* Plate 1337

``` {r,echo=FALSE}
per_1337
```

* Plate 1345

``` {r,echo=FALSE}
per_1345
```

* Plate 1350

``` {r,echo=FALSE}
per_1350
```

* Plate 1357

``` {r,echo=FALSE}
per_1357
```

* Plate 1360

``` {r,echo=FALSE}
per_1360
```

* Plate 1378

``` {r,echo=FALSE}
per_1378
```

* Plate 1385

``` {r,echo=FALSE}
per_1385
```

* Plate 1387

``` {r,echo=FALSE}
per_1387
```

```{r,echo=FALSE,eval=FALSE}
################################ Annotation ##################################

# here we select, from the EPIC characterizing data, the columns for:
# ILmnID, CHR, MAPINFO(geneposition), UCSC_RefGene_Name, USCS_RefGene_Group, CpG Island name, Phantom 4, Phantom5
selectedCols <- c("character", rep("NULL", 10), "character", "integer", rep("NULL", 2), "character", "NULL", rep("character", 2), "NULL", rep("character", 2), rep("NULL", 25))
EPICchar <- read.csv("EPIC MANIFEST AND SUPPORTING INFORMATION/MethylationEPIC_v-1-0_B1.csv", as.is=TRUE, colClasses = selectedCols)

# only choose rows with cpg sites in preprocessed data
DataChar <- EPICchar[EPICchar$IlmnID %in% row.names(noob_1337),]

## then add each sample
# data from noob preprocessed data
SampleData <- data.frame(getBeta(noob_1337), getBeta(noob_1345), getBeta(noob_1350), getBeta(noob_1357), getBeta(noob_1360), getBeta(noob_1378), getBeta(noob_1385), getBeta(noob_1387))
# rewrite sample names from patient file
sampleNames <- unlist(list(pat_1337$Sample_No, pat_1345$SAMPLE.ID, pat_1350$SAMPLE.ID, pat_1357$Sample.ID, pat_1360$Sample.ID, pat_1378$Tube.Label, pat_1385$Tube.Label, pat_1387$Sample.ID))
colnames(SampleData) <- sampleNames[-7]
# reorder sample data by cpg site (row name)
SampleData <- SampleData[order(row.names(SampleData)), ]

# save/load to save time
save(DataChar,file="myDC.Rdata")
save(SampleData,file="mySD.Rdata")
load("myDC.Rdata")
load("mySD.Rdata")

# merge two, then sort by IlmnID
FullAnnotation <- cbind(DataChar, SampleData)

save(FullAnnotation,file="myFA.Rdata")
load("myFA.Rdata")

```