### 2020-01-02 stan_Model.Rmd

---
title: "Stan Model"
output: word_document
---

This is the code for modeling our data using lmer function. The model we are using is:

$$A_i = b_i + \beta_T + b_T_i + \mu + \epsilon_i$$
$$B_i = b_i + \beta_T + b_T_i + \mu + \epsilon_i$$
$$N_i = b_i ~~~~~~~~~~~~~~~~~~+ \mu + \epsilon_i$$

* $A_i$, $B_i$, $N_i$ are the mythelation for side A, B, and N respectively
* $\beta_T$ = effect of tumor tissue (fixed)
* $\mu$ = grand mean
* $b_i \sim N(0,\sigma^2_P)$
* $b_T_i \sim N(0,\sigma^2_T)$
* $\epsilon_i \sim N(0,\sigma^2_e)$

Please download the "model.stan" file from Git and save it in the folder where all the other data are stored. Please also make sure the annotation file is saved in that folder.

This version of codes assigns uniform distribution as a prior for $\beta_T$, $\mu$, $\sigma^2_P$, $\sigma^2_T$, and $\sigma^2_y$.

``` {r, echo=FALSE, warning=FALSE, message=FALSE}
### install the following package
# install.package("rstan")

library(rstan)

# here set the working directory that points to the data folder
# e.g. the folder with annotated data saved as "myFA.Rdata"
# please just comment out the directories already here, so each
# user can uncomment the setwd for them as we move code around

# Kevin's working directory
# setwd("/Users/kevinmurgas/Documents/Data+ project/EPIC data")

# Yanlin's working directory
setwd("D:/DataPlus2017/Data")

# Set your directory here

### LOAD DATA

# load fully annotated data (saved from LoadDataAndQC.R)
load("myFA.Rdata")

# extract CpG site 1 to start
firstSite <- head(FullAnnotation, 1)

# convert first site to table, using only H E K* W C J (tumor+normal)
# 18 rows, with columns for patient, tumor/normal
# index in order A,B,N
indices <- c(11, 12, 13, 17, 18, 16, 20, 21, 19, 25, 26, 24, 33, 34, 32, 
             39, 31, 15)
patientLabel <- c(rep("H", 3),rep("E", 3),rep("K*", 3),rep("W", 3),rep("C", 3),
                  rep("J", 3))
tissueLabel <- rep(c("T", "T", "N"), 6)
sideLabel <- rep(c("A", "B", ""), 6)

# here we use all samples
indices <- c(9:13, 15:39,46,47,57,58,72:75)
patientLabel <- substr(colnames(firstSite[indices]),1,1)
patientLabel[10:12] <- "K*"
sideLabel <- substr(colnames(firstSite[indices]),2,2)
tissueLabel <- sideLabel
tissueLabel[tissueLabel %in% c("A","B")] <- "T"   #replace A and B with T
tissueIndicator <- 1*(tissueLabel=="T")

firstData <- data.frame(beta=t(firstSite[indices]), patient = patientLabel, tissue=tissueLabel, side=sideLabel, tInd=tissueIndicator)

### FIT WITH STAN

# Logit transform on beta values

firstData$y <- log(firstData$X1/(1-firstData$X1))

# Stan data

stanDat <- list(pID = as.integer(factor(firstData$patient)),
                tInd = firstData$tInd,
                N = nrow(firstData),
                P = nlevels(firstData$patient),
                y = firstData$y)

stanFit <- stan(file="model.stan",data=stanDat)
```

#### Summary of the Stan Model

``` {r, echo=FALSE,warning=FALSE}

print(stanFit, probs = c(0.025, 0.5, 0.975))

```

#### Estimatesfor the first three samples

``` {r, echo=FALSE,warning=FALSE}

est <- as.matrix(stanFit,pars=c("mu","betaT","b_pat","bT_pat"))

# Estimates for first three samples
head(est[,c(1:5,19:21)])
```

#### Graphs (still working on finding how useful these graphs are)

``` {r, echo=FALSE, warning=FALSE}
# Diagnostic Graphs

plot(stanFit, pars=c("b_pat"), main="Random Effect of Patients")
plot(stanFit, pars=c("bT_pat"), main="Random Effect of Tumor Tissue Varying Between Patients")

```