### 2020-09-25 StanAnalysis_updated.Rmd

---
title: "DNA Methylation Hierarchical Variance Model\nStan Results"
author: "Kevin Murgas"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: true
    number_section: true
    toc: true
params:
  prep: true
---

\newpage
\section{Model Overview}
\subsection{Data}
Logit-transformed beta values, from EPIC methylation array pre-processed with minfi Noob algorithm. \newline
\newline Sample counts (total): 48 samples from 21 patients. \newline
By tissue: 6 normal samples, 42 paired bulk tumor samples.
\subsection{Model equation}
Let i=sample, j=patient
$$ y_i = \mu + a_j +  \text{tInd}(i)*(\beta_T + b_j + c_i) $$
$$ a_j \sim N(0,\sigma_P) , b_j \sim N(0,\sigma_{PT}) , c_i \sim N(0,\sigma_T) $$
\newline Model is fit at each individual CpG site using RStan's MCMC sampling algorithm. \newline
Stan Parameters: 4 chains x 2000 iterations (200 warmup) = 7200 total draws. adapt_delta = 0.999

\subsection{Prior Probabilities}
Priors for each parameter (mu, betaT, sigmaP/PT/T) were defined using TCGA 450k array data from colorectal tumor patients with paired multiple tumor and/or normal+tumor samples. \newline Empirical values of mean and standard deviations were assesed for all 450k CpG sites. \newline The resulting distributions were fit with: bimodal gaussian (mu), Cauchy (betaT), or gamma distributions (sigmas). \newline
SigmaE (error variance) prior was defined using a non-informative gamma distribution
\newline Additionally: Prior relaxation was performed for the Cauchy and gamma distribtions by algebraically increasing the variance of the distributions while maintaining the same mode (peak)

\subsection{Conservation score}
We want to understand which sites of DNA methylation are fundamentally conserved within tumors, i.e. lower variation in methlyation within tumor (sigmaT) relative to the variation within normal healthy tissue (sigmaP). Therefore we choose the log-ratio of sigmaP/sigmaT to represent a conservation score, which is positive when tumor variation is lower than normal variation.
We take the median of the posterior distribution for this parameter. Additionally, we calculate a regularization term based on the inverse of the L2 norm of sigmaP and sigmaT.

<!--  $$ \text{score}_\text{prob} = \Pr (\text{log}\frac{\sigma_{P}}{\sigma_T} > 0) = \int \text{log}\frac{\sigma_{P}}{\sigma_T} > 0 ~ dp $$
$$ \text{score}_\text{mean} = \text{Mean} (\text{log}\frac{\sigma_{P}}{\sigma_T}) ,    ~ -->
$$\text{score}_\text{med} = \text{Median} (\text{log}\frac{\sigma_{P}}{\sigma_T}) $$

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gt)
library(rlang)
library(ggpubr)
library(knitr)

# set working directory
setwd("/Users/kevinmurgas/Documents/Data+ project/EPIC data")
library(here)

select <- dplyr::select
rename <- dplyr::rename
filter <- dplyr::filter

set.seed(1234)
```

```{r Load, include=FALSE}
# Load raw methylation data
load(here("myFA_bulkonly.Rdata"))

stanFile = "FullResultsTCGA_relaxgamma3_postsum3.Rdata"
load(here(stanFile))
```

\newpage
\section{Fit Results}
\subsection{Parameter Distributions}
Histograms are median values of posterior distribution
```{r Params_site, echo=FALSE, fig.fullwidth=TRUE, warning=FALSE, comment=NA}
genemask = !(FullAnnotation$UCSC_RefGene_Name=="")
cat(sprintf("N sites: total=%i, gene-associated=%i, non-gene=%i  \n",length(genemask),sum(genemask),length(genemask)-sum(genemask)))

fBG <- function(x, a, m1, s1, m2, s2) a*dnorm(x,m1,s1) + (1-a)*dnorm(x,m2,s2) # bimodal gaussian
post_prior_histogram <- function(postvals,mask,priorfun,bw,xl,titlestr) {
  # stacked histogram colored by gene-associated sites
  # and prior curve overlaid
  plot_df <- data.frame(val=postvals, gene_assoc=mask)
  p1 <- ggplot(plot_df, aes(x=val)) +
    geom_histogram(aes(y=..count../length(mask),fill=gene_assoc), binwidth=bw) +
    stat_function(fun=function(x) bw*priorfun(x),geom="line",color="black") +
    ggtitle(titlestr) + xlim(xl) + theme_classic() + theme(legend.position = "none") + xlab(titlestr) + ylab("Freq.")
  p1
}

# plot mu,betaT,sigmaE,sigmaP,sigmaPT,sigmaT with priors overlaid
pM <- post_prior_histogram(mu_full$p50, genemask, function(x) 1.5*fBG(x,0.3557,-3.0675, 0.75214,1.1290, 1.39197), 0.1, c(-5,5), "Mu")
pB <- post_prior_histogram(betaT_full$p50, genemask, function(x) 1.7*dcauchy(x,-0.04051, 0.5799), 0.1, c(-5,5), "BetaT")
pE <- post_prior_histogram(sigmaE_full$p50, genemask, function(x) 4*dgamma(x,2,2), 0.05, c(0,2), "SigmaE")
pP <- post_prior_histogram(sigmaP_full$p50, genemask, function(x) 2*dgamma(x, 2.7693, 7.0029), 0.05, c(0,2), "SigmaP")
pPT <- post_prior_histogram(sigmaPT_full$p50, genemask, function(x) 1.8*dgamma(x, 2.1457, 2.4819), 0.05, c(0,2), "SigmaPT")
pT <- post_prior_histogram(sigmaT_full$p50, genemask, function(x) 1.2*dgamma(x, 3.3643, 10.430), 0.05, c(0,2), "SigmaT")
# log sigmas
pLP <- post_prior_histogram(log(sigmaP_full$p50), genemask, function(x) 0.52*dgamma(exp(x), 2.7693, 7.0029), 0.1, c(-5,5), "log(SigmaP)")
pLPT <- post_prior_histogram(log(sigmaPT_full$p50), genemask, function(x) 0.7*dgamma(exp(x), 2.1457, 2.4819), 0.1, c(-5,5), "log(SigmaPT)")
pLT <- post_prior_histogram(log(sigmaT_full$p50), genemask, function(x) 0.4*dgamma(exp(x), 3.3643, 10.430), 0.1, c(-5,5), "log(SigmaT)") + theme(legend.position = c(0.8, 0.6))

# arrange each parameter as a subplot # ,labels=c("M","B","E","P","PT","T","lP","lPT","lT")
p_all <- ggarrange(pM,pB,pE,pP,pPT,pT,pLP,pLPT,pLT,hjust=0,nrow=3,ncol=3)
print(p_all)
rm(pM,pB,pE,pP,pPT,pT,pLP,pLPT,pLT)
```

\subsection{Conservation scores}
```{r Fig1B, echo=FALSE, fig.height=2, fig.width=6.5, fig.fullwidth=TRUE, warning=FALSE}

post_histogram <- function(postvals,mask,bw,xl,titlestr) {
  # stacked histogram colored by gene-associated sites
  # and prior curve overlaid
  plot_df <- data.frame(val=postvals, gene_assoc=mask)
  p1 <- ggplot(plot_df, aes(x=val)) +
    geom_histogram(aes(y=..count../length(mask),fill=gene_assoc), binwidth=bw) +
    ggtitle(titlestr) + xlim(xl) + theme_classic() + theme(legend.position = "none") + xlab(titlestr) + ylab("Freq.")
}

# plot the two conservations score prob sums
pP1 <- post_histogram(prob1_full$median,genemask,0.05,c(-3,3),expression("Median log("~sigma[P]/sigma[T]~")"))
prob1_proxy = log(sigmaP_full$p50/sigmaT_full$p50)
pP2 <- post_histogram(prob1_proxy,genemask,0.05,c(-3,3),expression(" log( median" ~sigma[P]/~"median"~sigma[T]~")"))
# regularized version
L2norm = sqrt(sigmaP_full$p50^2 + sigmaT_full$p50^2)
prob1_reg = log(sigmaP_full$p50/sigmaT_full$p50) - 1/L2norm
pP3 <- post_histogram(prob1_reg,genemask,0.1,c(-10,5),"Regularized")
p_probs <- ggarrange(pP1,pP2,pP3,nrow=1,ncol=3)
print(p_probs)
```

\newpage
\subsection{Summary of neff and Rhat for each parameter}
```{r neffRhat, echo=FALSE, warning=FALSE}
# for each parameter (mu, betaT, sigmaP, sigmaPT, sigmaT, sigmaE, lp__)
# reporting: mean n_eff, min n_eff, # n_eff<500, %, # n_eff<100, %
n_eff_mat <-
  data.frame(
  mu = mu_full$n_eff,
  betaT = betaT_full$n_eff,
  sigmaP = sigmaP_full$n_eff,
  sigmaPT = sigmaPT_full$n_eff,
  sigmaT = sigmaT_full$n_eff,
  sigmaE = sigmaE_full$n_eff,
  lp = lp_full$n_eff
  ) %>% gather(param,value,mu:lp) %>% group_by(param) %>% summarize(
  mean = mean(value),
  min = min(value),
  s500 = sum(value < 500),
  p500 = 100 * mean(value < 500),
  s100 = sum(value < 100),
  p100 = 100 * mean(value < 100),
  .groups = "drop"
  ) %>% arrange(match(.$param,c("mu","betaT","sigmaP","sigmaPT","sigmaT","sigmaE","lp")))

kable(n_eff_mat, col.names=c("Parameter", "Mean n_eff", "Min n_eff", "# <500", "% <500", "# <100", "% <100"), digits = 2)

# again for Rhat
Rhat_mat <-
  data.frame(
  mu = mu_full$Rhat,
  betaT = betaT_full$Rhat,
  sigmaP = sigmaP_full$Rhat,
  sigmaPT = sigmaPT_full$Rhat,
  sigmaT = sigmaT_full$Rhat,
  sigmaE = sigmaE_full$Rhat,
  lp = lp_full$Rhat
  ) %>% gather(param,value,mu:lp) %>% group_by(param) %>% summarize(
  mean = mean(value),
  max = max(value),
  s11 = sum(value > 1.1),
  p11 = 100 * mean(value > 1.1),
  s101 = sum(value > 1.01),
  p101 = 100 * mean(value > 1.01),
  .groups = "drop"
  ) %>% arrange(match(.$param,c("mu","betaT","sigmaP","sigmaPT","sigmaT","sigmaE","lp")))

kable(Rhat_mat, col.names=c("Parameter", "Mean Rhat", "Max Rhat", "# >1.1", "% >1.1", "# >1.01", "% >1.01"), digits = 2)
cat("  \n")

# indices to filter out sites Rhat>1,1
QCinds <- lp_full$Rhat > 1.1
```

For the remainder of these results, a filter is applied to remove all sites with Rhat > 1.1 in lp (log-posterior likelihood of entire site fit). Another filter is later applied to remove all genes with <5 CpG sites.

\newpage Prep data
```{r prep, echo=FALSE, include=TRUE, eval=params$prep, comment=NA}
cat("Param:prep=TRUE -> prepare data...")

med_vals <- tibble(Name=FullAnnotation[,1], mu=mu_full$p50, betaT=betaT_full$p50, sigmaE=sigmaE_full$p50, sigmaP=sigmaP_full$p50, sigmaPT=sigmaPT_full$p50, sigmaT=sigmaT_full$p50, Rhat=lp_full$Rhat) %>% mutate(logP=log(sigmaP), logPT=log(sigmaPT), logT=log(sigmaT)) %>%
  mutate(L2norm = sqrt(sigmaP^2 + sigmaT^2)) %>%
  mutate(score_median=prob1_full$median, score_median2=log(sigmaP/sigmaT), score_reg=log(sigmaP/sigmaT * L2norm))
# score_prob=prob1_full$prob, score_mean=prob1_full$mean,

# define filter cutoffs
Rhat_cutoff=1.1 # maximum Rhat cutoff
n_cutoff=5 # cutoff for gene minimum number of CpG sites

# prep data by joining with FullAnnotation (columns: IlmnID, CHR, MAPINFO, RefGene_Name+Group, CpG_Island)
cat("Prep data with annotation, filter out sites Rhat>1.1 (and save)")
ptm<-proc.time()
prepped_data <- tibble(FullAnnotation[,c(1,2,3,5,6,8)]) %>% rename(Name = IlmnID) %>%
  left_join(by="Name", med_vals) %>% filter(Rhat<Rhat_cutoff)
write_csv(prepped_data, here("data-stan_rg3ps3", "prepped_data.csv"))
cat(proc.time()-ptm)

# separate by rows by genes
cat("Separate by gene")
ptm<-proc.time()
gened_data <- prepped_data %>%
  arrange(CHR, MAPINFO) %>%
  separate_rows(UCSC_RefGene_Name, sep=";") %>% 
  distinct()
write_csv(gened_data, here("data-stan_rg3ps3", "gened_data.csv"))
cat(proc.time()-ptm)

gene_summarize <- function(data, id, value, fun = mean, ...) {
  grouped_data <- data %>% group_by({{id}})

  bind_cols(
    summarise_at(grouped_data, dplyr::vars({{value}}), fun, ...),
    count(grouped_data) %>% dplyr::ungroup() %>% dplyr::select(.data$n)
  )
}

cat(paste0("Summarize by gene, taking mean value of all sites in gene, filter out sites n<",n_cutoff))
ptm<-proc.time()
gene_summary <- gened_data %>%
  filter(!is.na(UCSC_RefGene_Name)) %>%
  gene_summarize(UCSC_RefGene_Name, c(mu:sigmaT,score_median:score_reg)) %>% filter(n>=n_cutoff)
cat(proc.time()-ptm)

boot_naive <- function(data, data_full, value, id, reps) {
  lengths <- sort(unique(data$n))
  varname <- paste0(as_name(ensym(value)), "_v1")

  # perm_v1_inner fxn: draw n_rep samples of n_sites CpG sites, randomly out of all sites
  # take mean of each draw
  sample_naive <- function(n_sites, n_rep, data) {
    replicate(n = n_rep, expr = mean(sample(data, as.numeric(n_sites), TRUE)))
  }

  # run naive_sample on all lengths to get res (list of null distributions for each length)
  # then map a percentile to each gene based on null distrib. of that length
  names(lengths) <- lengths
  res <- lapply(lengths, sample_naive, n_rep = reps, data = pull(data_full, {{value}}))
  values <- map2_dbl(pull(data, {{value}}), data$n, ~ mean(res[[as.character(.y)]] < .x))

  data %>% mutate(!!varname := values)
}

boot_adj <- function(data, data_full, value, id, reps) {
  lengths <- sort(unique(data$n))
  varname <- paste0(as_name(ensym(value)), "_v3")

  sample_adj <- function(n_sites, n_rep, data, id, value) {
    perm3_data <- data %>% group_by({{id}}) %>% mutate(left = n() - row_number() + 1)
    starting_index <- sample(which(n_sites <= perm3_data$left), n_rep, replace = TRUE)
    map_dbl(starting_index, ~ mean(pull(perm3_data, {{value}})[.x + seq_len(n_sites) - 1]))
  }

  names(lengths) <- lengths
  res <- lapply(lengths, sample_adj, n_rep = reps, data = data_full, id = {{id}}, value = {{value}})
  values <- map2_dbl(pull(data, {{value}}), data$n, ~ mean(res[[as.character(.y)]] < .x))

  data %>% mutate(!!varname := values)
}

cat("Permutate data with bootstrapping")
ptm<-proc.time()
# 2nd function of methcon5
permed_data <- gene_summary %>%
  filter(n < quantile(n, 0.99)) %>%
  boot_naive(gened_data, sigmaP, UCSC_RefGene_Name, 1000) %>%
  boot_naive(gened_data, sigmaPT, UCSC_RefGene_Name, 1000) %>%
  boot_naive(gened_data, sigmaT, UCSC_RefGene_Name, 1000) %>%
  boot_naive(gened_data, score_median, UCSC_RefGene_Name, 1000) %>%
  boot_naive(gened_data, score_median2, UCSC_RefGene_Name, 1000) %>%
  boot_naive(gened_data, score_reg, UCSC_RefGene_Name, 1000) %>%
  boot_adj(gened_data, sigmaP, UCSC_RefGene_Name, 1000) %>%
  boot_adj(gened_data, sigmaPT, UCSC_RefGene_Name, 1000) %>%
  boot_adj(gened_data, sigmaT, UCSC_RefGene_Name, 1000) %>%
  boot_adj(gened_data, score_median, UCSC_RefGene_Name, 1000) %>%
  boot_adj(gened_data, score_median2, UCSC_RefGene_Name, 1000) %>%
  boot_adj(gened_data, score_reg, UCSC_RefGene_Name, 1000)
write_csv(permed_data, here("data-stan_rg3ps3", "permed_data.csv"))
cat(proc.time()-ptm)
#  boot_naive(gened_data, score_prob, UCSC_RefGene_Name, 1000) %>%
#  boot_naive(gened_data, score_mean, UCSC_RefGene_Name, 1000) %>%

# look at mean methylation conservation in promoters
cat("Calculate mean methylation conservation in promoters")
ptm<-proc.time()
promoted_data <- prepped_data %>%
  mutate(gene = str_extract(UCSC_RefGene_Name, "^[^;]+")) %>%
  filter(!is.na(gene)) %>%
  left_join(read_csv(here("data-raw","MethylationEPIC_v-1-0_B4.csv"), skip = 7) %>%  
              select(Name, Regulatory_Feature_Group), 
            by = "Name") %>%
  filter(Regulatory_Feature_Group == "Promoter_Associated") %>%
  group_by(gene) %>%
  summarise(sigmaP_pro = mean(sigmaP), sigmaPT_pro = mean(sigmaPT), sigmaT_pro = mean(sigmaT), 
            score_median_pro = mean(score_median), score_median2_pro = mean(score_median2), score_reg_pro = mean(score_reg))
write_csv(promoted_data, here("data-stan_rg3ps3", "promoted_data.csv"))
cat(proc.time()-ptm)
cat("Done with data preparation (BOTH)")
```

```{r Exit, include=FALSE}
knit_exit()
```

```{r load_prep, echo=FALSE, include=TRUE, message=FALSE, eval=!params$prep}
cat("Param:prep=FALSE -> load saved data (BOTH)...")
gened_data <- read_csv(here("data-stan_rg3ps3", "gened_data.csv"), col_types = list("CHR" = col_character()))
prepped_data <- read_csv(here("data-stan_rg3ps3", "prepped_data.csv"))
permed_data <- read_csv(here("data-stan_rg3ps3", "permed_data.csv"))
promoted_data <- read_csv(here("data-stan_rg3ps3", "promoted_data.csv"))
cat("done")
```

``` {r enrich, include=FALSE}
# expression data
# downloaded file: E-MTAB-2836-query-results.tpms.tsv
# source: https://www.ebi.ac.uk/gxa/experiments/E-MTAB-2836/Results
expression_atlas <- read_tsv(here("data-raw", "E-MTAB-2836-query-results.tpms.tsv"), skip = 4, 
                             col_types = cols(
                               .default = col_double(),
                               `Gene ID` = col_character(),
                               `Gene Name` = col_character()
                             ))

# Gene Set Enrichment
tidy_gsea <- function(genes) {
  AnnotationDbi::mapIds(org.Hs.eg.db::org.Hs.eg.db, genes, 'ENTREZID', 'SYMBOL') %>%
    ReactomePA::enrichPathway()
}

enrich_sigmaP <- permed_data %>%
  #filter(sigmaP_v3 < quantile(sigmaP_v3, 0.05, na.rm = TRUE)) %>%
  filter(sigmaP_v3 < 0.05) %>%
  pull(UCSC_RefGene_Name) %>% tidy_gsea()
enrich_sigmaPT <- permed_data %>%
  #filter(sigmaPT_v3 < quantile(sigmaPT_v3, 0.05, na.rm = TRUE)) %>%
  filter(sigmaPT_v3 < 0.05) %>%
  pull(UCSC_RefGene_Name) %>% tidy_gsea()
enrich_sigmaT <- permed_data %>%
  #filter(sigmaT_v3 < quantile(sigmaT_v3, 0.05, na.rm = TRUE)) %>%
  filter(sigmaT_v3 < 0.05) %>%
  pull(UCSC_RefGene_Name) %>% tidy_gsea()
#enrich_score_prob <- permed_data %>%
#  filter(score_prob_v3 > quantile(score_prob_v3, 0.95, na.rm = TRUE)) %>%
#  pull(UCSC_RefGene_Name) %>% tidy_gsea()
#enrich_score_mean <- permed_data %>%
#  filter(score_mean_v3 > quantile(score_mean_v3, 0.95, na.rm = TRUE)) %>%
#  pull(UCSC_RefGene_Name) %>% tidy_gsea()
enrich_score_median <- permed_data %>%
  #filter(score_median_v3 > quantile(score_median_v3, 0.95, na.rm = TRUE)) %>%
  filter(score_median_v3 > 0.95) %>%
  pull(UCSC_RefGene_Name) %>% tidy_gsea()
enrich_score_median2 <- permed_data %>%
  #filter(score_median2_v3 > quantile(score_median2_v3, 0.95, na.rm = TRUE)) %>%
  filter(score_median2_v3 > 0.95) %>%
  pull(UCSC_RefGene_Name) %>% tidy_gsea()
enrich_score_reg <- permed_data %>%
  #filter(score_reg_v3 > quantile(score_reg_v3, 0.95, na.rm = TRUE)) %>%
  filter(score_reg_v3 > 0.95) %>%
  pull(UCSC_RefGene_Name) %>% tidy_gsea()
```

\newpage
\subsection{Tables comparing different groups of CpG sites:}
```{r Table1, echo=FALSE, warning=FALSE}
#, include=TRUE, results='asis', fig.keep='all', fig.show='asis', fig.height=3, fig.width=5
# Table 1
prepped_data %>%
  group_by(`Gene Associated` = !is.na(UCSC_RefGene_Name)) %>%
  summarise_at(vars(sigmaP:sigmaT,score_median:score_reg), mean, na.rm = TRUE) %>%
  mutate(`Gene Associated` = factor(`Gene Associated`, c(TRUE, FALSE), c("Yes", "No"))) %>%
  arrange(rev(rownames(.))) %>%
  gt() %>%
  fmt_number(columns = vars(sigmaP, sigmaPT, sigmaT, score_median, score_median2, score_reg), decimals = 3) %>%
  tab_spanner(label = "Hierarchical Variance", columns = vars(sigmaP, sigmaPT, sigmaT)) %>%
  tab_spanner(label = "Conservation Scores", columns = vars(score_median, score_median2, score_reg))
#score_prob, score_mean,

prepped_data %>%
  group_by(`Island relation` = Relation_to_UCSC_CpG_Island) %>%
  summarise_at(vars(sigmaP:sigmaT,score_median:score_reg), mean, na.rm = TRUE) %>%
  dplyr::rename(ir = `Island relation`) %>%
  mutate(ir = case_when(is.na(ir) ~ "Sea",
                        str_detect(ir, "^S") ~ str_replace(ir, "^S_", "South "),
                        str_detect(ir, "^N") ~ str_replace(ir, "^N_", "North "),
                        TRUE ~ ir)) %>%
  dplyr::rename(`Island relation` = ir) %>%
  arrange(rev(rownames(.))) %>%
  gt() %>%
  fmt_number(columns = vars(sigmaP, sigmaPT, sigmaT, score_median, score_median2, score_reg), decimals = 3) %>%
  tab_spanner(label = "Hierarchical Variance", columns = vars(sigmaP, sigmaPT, sigmaT)) %>%
  tab_spanner(label = "Conservation Scores", columns = vars(score_median, score_median2, score_reg))

prepped_data %>%
  group_by(`5'UTR` = str_detect(UCSC_RefGene_Group, "5'UTR") & !is.na(UCSC_RefGene_Group)) %>%
  summarise_at(vars(sigmaP:sigmaT,score_median:score_reg), mean, na.rm = TRUE) %>%
  mutate(`5'UTR` = factor(`5'UTR`, c(TRUE, FALSE), c("Yes", "No"))) %>%
  arrange(rev(rownames(.))) %>%
  gt() %>%
  fmt_number(columns = vars(sigmaP, sigmaPT, sigmaT, score_median, score_median2, score_reg), decimals = 3) %>%
  tab_spanner(label = "Hierarchical Variance", columns = vars(sigmaP, sigmaPT, sigmaT)) %>%
  tab_spanner(label = "Conservation Scores", columns = vars(score_median, score_median2, score_reg))

prepped_data %>%
  group_by(`TSS1500` = str_detect(UCSC_RefGene_Group, "TSS1500") & !is.na(UCSC_RefGene_Group)) %>%
  summarise_at(vars(sigmaP:sigmaT,score_median:score_reg), mean, na.rm = TRUE) %>%
  mutate(`TSS1500` = factor(`TSS1500`, c(TRUE, FALSE), c("Yes", "No"))) %>%
  arrange(rev(rownames(.))) %>%
  gt() %>%
  fmt_number(columns = vars(sigmaP, sigmaPT, sigmaT, score_median, score_median2, score_reg), decimals = 3) %>%
  tab_spanner(label = "Hierarchical Variance", columns = vars(sigmaP, sigmaPT, sigmaT)) %>%
  tab_spanner(label = "Conservation Scores", columns = vars(score_median, score_median2, score_reg))

prepped_data %>%
  group_by(`TSS200` = str_detect(UCSC_RefGene_Group, "TSS200") & !is.na(UCSC_RefGene_Group)) %>%
  summarise_at(vars(sigmaP:sigmaT,score_median:score_reg), mean, na.rm = TRUE) %>%
  mutate(`TSS200` = factor(`TSS200`, c(TRUE, FALSE), c("Yes", "No"))) %>%
  arrange(rev(rownames(.))) %>%
  gt() %>%
  fmt_number(columns = vars(sigmaP, sigmaPT, sigmaT, score_median, score_median2, score_reg), decimals = 3) %>%
  tab_spanner(label = "Hierarchical Variance", columns = vars(sigmaP, sigmaPT, sigmaT)) %>%
  tab_spanner(label = "Conservation Scores", columns = vars(score_median, score_median2, score_reg))
```

\newpage
\subsection{Gene-average Parameter Distributions}
```{r Params_gene, echo=FALSE, fig.fullwidth=TRUE, warning=FALSE, comment=NA}
# make gene-wise averages
gene_summarize <- function(data, id, value, fun = mean, ...) {
  grouped_data <- data %>% group_by({{id}})

  bind_cols(
    summarise_at(grouped_data, dplyr::vars({{value}}), fun, ...),
    count(grouped_data) %>% dplyr::ungroup() %>% dplyr::select(.data$n)
  )
}

cat("Summarize by gene, taking mean value of all sites in gene")
ptm<-proc.time()
gene_summary <- gened_data %>%
  filter(!is.na(UCSC_RefGene_Name)) %>%
  gene_summarize(UCSC_RefGene_Name, c(mu:sigmaT,score_median:score_reg))
cat(proc.time()-ptm)

post_histogram <- function(postvals,bw,xl,titlestr) {
  # stacked histogram colored by gene averages
  # and prior curve overlaid
  plot_df <- data.frame(val=postvals)
  p1 <- ggplot(plot_df, aes(x=val)) +
    geom_histogram(aes(y=..density..), binwidth=bw) +
    ggtitle(titlestr) + xlim(xl) + theme_classic() + xlab(titlestr) + ylab("Freq.")
  p1
}

# plot: n_sites, mu, betaT, sigmaP, sigmaPT, sigmaT
n_cutlo = 5
n_cuthi = quantile(gene_summary$n, 0.99)
pN <- post_histogram(gene_summary$n,1,c(0,200),paste0("N sites (max = ",max(gene_summary$n),")")) + geom_vline(xintercept=n_cutlo-0.5,linetype = "dashed",color="red") + annotate(geom="text",x=80,y=0.03,size=2,label=paste0("# genes <",n_cutlo," sites:\n", sum(gene_summary$n<n_cutlo)," / ",nrow(gene_summary))) +
geom_vline(xintercept=n_cuthi,linetype = "dashed",color="red") +annotate(geom="text",x=120,y=0.013,size=2,label=sprintf("# genes >%0.1f sites:\n %i / %i (99th pctile)", n_cuthi,sum(gene_summary$n>n_cuthi),nrow(gene_summary)))
gene_summary <- gene_summary %>% filter(n>=n_cutlo) %>% filter(n<n_cuthi)
pM <- post_histogram(gene_summary$mu,0.1,c(-5,5),"Mu")
pB <- post_histogram(gene_summary$betaT,0.1,c(-5,5),"BetaT")
pP <- post_histogram(gene_summary$sigmaP,0.05,c(0,2),"SigmaP")
pPT <- post_histogram(gene_summary$sigmaPT,0.05,c(0,2),"SigmaPT")
pT <- post_histogram(gene_summary$sigmaT,0.05,c(0,2),"SigmaT")
pP1 <- post_histogram(gene_summary$score_median,0.05,c(-3,3),"Score_median")
pP2 <- post_histogram(gene_summary$score_median2,0.05,c(-3,3),"Score_median2")
pP3 <- post_histogram(gene_summary$score_reg,0.05,c(-3,3),"Score_reg")

#pP1 <- post_histogram(gene_summary$score_prob,0.01,c(0,1),"Score_prob")
#pP2 <- post_histogram(gene_summary$score_mean,0.05,c(-3,3),"Score_mean")

# arrange each parameter as a subplot
p_all <- ggarrange(pN,pM,pB,pP,pPT,pT,pP1,pP2,pP3,hjust=0,nrow=3,ncol=3)
print(p_all)
rm(pN,pM,pB,pP,pPT,pT,pP1,pP2,pP3)

```

\newpage
\subsection{Gene-average parameters vs N sites}
```{r Params_vsN, echo=FALSE, fig.fullwidth=TRUE, warning=FALSE, comment=NA, dev = "png"}
pM <- ggplot(gene_summary) + geom_point(aes(x=n,y=mu), alpha=0.01) + ggtitle("Mu vs N-sites")
pB <- ggplot(gene_summary) + geom_point(aes(x=n,y=betaT), alpha=0.01) + ggtitle("BetaT vs N-sites")
pE <- ggplot(gene_summary) + geom_point(aes(x=n,y=sigmaE), alpha=0.01) + ggtitle("SigmaE vs N-sites")
pP <- ggplot(gene_summary) + geom_point(aes(x=n,y=sigmaP), alpha=0.01) + ggtitle("SigmaP vs N-sites")
pPT <- ggplot(gene_summary) + geom_point(aes(x=n,y=sigmaPT), alpha=0.01) + ggtitle("SigmaPT vs N-sites")
pT <- ggplot(gene_summary) + geom_point(aes(x=n,y=sigmaT), alpha=0.01) + ggtitle("SigmaT vs N-sites")

#pP1 <- ggplot(gene_summary) + geom_point(aes(x=n,y=score_prob), alpha=0.01) + ggtitle("Score_prob vs N-sites")
#pP2 <- ggplot(gene_summary) + geom_point(aes(x=n,y=score_mean), alpha=0.01) + ggtitle("Score_mean vs N-sites")
pP1 <- ggplot(gene_summary) + geom_point(aes(x=n,y=score_median), alpha=0.01) + ggtitle("Score_median vs N-sites")
pP2 <- ggplot(gene_summary) + geom_point(aes(x=n,y=score_median2), alpha=0.01) + ggtitle("Score_median2 vs N-sites")
pP3 <- ggplot(gene_summary) + geom_point(aes(x=n,y=score_reg), alpha=0.01) + ggtitle("Score_reg vs N-sites")

# arrange each parameter as a subplot
p_all <- ggarrange(pM,pB,pE,pP,pPT,pT,pP1,pP2,pP3,hjust=0,nrow=3,ncol=3)
print(p_all)
rm(pM,pB,pE,pP,pPT,pT,pP1,pP2,pP3)
```

\newpage
\subsection{Figure 1: CpG Distance Analysis}
Corrected for gene direction, based on Ensembl gene database
```{r Fig1, echo=FALSE, warning=FALSE, fig.width=7, fig.height=2.5, comment=NA}
# load ensembl db
library(EnsDb.Hsapiens.v75) # ensembl annotation with gene position (hg19 version)
edb <- EnsDb.Hsapiens.v75
geneKeys <- keys(edb, keytype="GENENAME") %>% intersect(unique(gened_data$UCSC_RefGene_Name))
edbData<- AnnotationDbi::select(edb, keys=geneKeys, keytype="GENENAME",
                                columns=c("GENEBIOTYPE", "GENESEQSTART", "GENESEQEND", "SEQNAME", "SEQSTRAND")) %>%
  filter(SEQNAME %in% c(as.character(1:22),"X","Y")) %>%
  mutate(GENESTART = GENESEQSTART*(SEQSTRAND==1) + GENESEQEND*(SEQSTRAND==-1)) %>%
  select(-c(GENESEQSTART,GENESEQEND))

# calculate CpG distance based on Ensembl gene start position and direction
distanced_data <- gened_data %>%
  filter(!is.na(UCSC_RefGene_Name)) %>%
  filter(UCSC_RefGene_Name %in% geneKeys) %>%
  mutate(EDBCHR=edbData$SEQNAME[match(UCSC_RefGene_Name,edbData$GENENAME)],
         EDBSTART=edbData$GENESTART[match(UCSC_RefGene_Name,edbData$GENENAME)],
         EDBDIR=edbData$SEQSTRAND[match(UCSC_RefGene_Name,edbData$GENENAME)],
         EDBTYPE=edbData$GENEBIOTYPE[match(UCSC_RefGene_Name,edbData$GENENAME)]
         ) %>%
  filter(CHR==EDBCHR) %>%
  mutate(cpgdist = if_else(EDBDIR==1, MAPINFO-EDBSTART, EDBSTART-MAPINFO))

distance_length <- distanced_data %>%
  filter(!is.na(UCSC_RefGene_Name)) %>%
  select(CHR, UCSC_RefGene_Name, cpgdist, c(sigmaP:sigmaT,score_median:score_reg)) %>%
  mutate(location_bin = cut(cpgdist, c(seq(-1500, 20000, by = 100)), include.lowest = TRUE)) %>%
  group_by(location_bin) %>%
  summarise(mean_sigmaP = mean(sigmaP), mean_sigmaPT = mean(sigmaPT), mean_sigmaT = mean(sigmaT),  mean_score_median = mean(score_median), mean_score_median2 = mean(score_median2), n_obs = n(), .groups='drop') %>%
  mutate(location_bin = as.numeric(str_extract(location_bin, "(?<=,)(.*)(?=])")))

distance_length %>%
  drop_na() %>%
  pivot_longer(mean_sigmaP:mean_sigmaT) %>% 
  mutate(name = factor(name, c("mean_sigmaP","mean_sigmaPT","mean_sigmaT"), c("SigmaP","SigmaPT","SigmaT"))) %>%
  sample_frac() %>%
  ggplot(aes(location_bin, value, color = name, fill = name)) +
  geom_point(key_glyph = draw_key_rect) +
  ggforce::facet_zoom(location_bin < 5000) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(title = NULL,
       fill = NULL,
       color = NULL,
       x = "Distance from gene start site (binned to 100 bp)",
       y = "Average Sigma Value", 
       caption = "Figure 1: Average Sigma values of single CpGs as a function of position relative to gene.") +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2")
cat("\n")

# again for conservation scores
distance_length %>%
  drop_na() %>%
  pivot_longer(mean_score_median:mean_score_median2) %>% 
  mutate(name = factor(name, c("mean_score_median","mean_score_median2"), c("Score_median","Score_median2"))) %>%
  sample_frac() %>%
  ggplot(aes(location_bin, value, color = name, fill = name)) +
  geom_point(key_glyph = draw_key_rect) +
  ggforce::facet_zoom(location_bin < 5000) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(title = NULL,
       fill = NULL,
       color = NULL,
       x = "Distance from gene start site (binned to 100 bp)",
       y = "Average Conservation Score", 
       caption = "Figure 1: Average conservation scores of single CpGs as a function of position relative to gene.") +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2")
cat("\n")

# and look at number of CpG in each bin
distance_length %>%
  drop_na() %>%
  pivot_longer(n_obs) %>%
  mutate(name = factor(name, c("n_obs"), c("n_obs"))) %>%
  sample_frac() %>%
  ggplot(aes(location_bin, value, color = name, fill = name)) +
  geom_point(key_glyph = draw_key_rect) +
  ggforce::facet_zoom(location_bin < 5000) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(title = NULL,
       fill = NULL,
       color = NULL,
       x = "Distance from gene start site (binned to 100 bp)",
       y = "Number of CpGs", 
       caption = "Figure 1: Frequency of single CpGs as a function of position relative to gene start site.") +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2")
```

\subsection{Figure 2: Bootstrapping}
```{r Fig2, echo=FALSE, warning=FALSE, fig.height=4, comment=NA}
# Figure 2
permed_data %>%
  pivot_longer(dplyr::matches("1|3")) %>%
  separate(name, into = c("type", "method"), sep = "_v") %>%
  mutate(method = factor(method, c("1", "3"), c("Naive bootstrap", "Adjusted bootstrap")),
         type = factor(type, c("sigmaP","sigmaPT","sigmaT"), c("SigmaP","SigmaPT","SigmaT"))) %>%
  drop_na() %>%
  ggplot(aes(value, fill = type)) +
  geom_histogram(bins = 50, color = "grey30") +
  facet_grid(method~type, scales = "free") +
  theme_minimal() +
  labs(x = "p-value",
       y = "Number of genes",
       title = NULL,
       caption = "Figure 2A: Distribution of boot-strapping p-values for genes as a function of sigma level and bootstrap method.")  +
  theme(legend.position = "none") +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2")

cat("***")
permed_data %>%
  pivot_longer(dplyr::matches("1|3")) %>%
  separate(name, into = c("type", "method"), sep = "_v") %>%
  mutate(method = factor(method, c("1", "3"), c("Naive bootstrap", "Adjusted bootstrap")),
         type = factor(type, c("score_median","score_median2","score_reg"), c("Score_median","Score_median2","Score_reg"))) %>%
  drop_na() %>%
  ggplot(aes(value, fill = type)) +
  geom_histogram(bins = 50, color = "grey30") +
  facet_grid(method~type, scales = "free") +
  theme_minimal() +
  labs(x = "p-value",
       y = "Number of genes",
       title = NULL,
       caption = "Figure 2B: Distribution of boot-strapping p-values for genes as a function of conservation score and bootstrap method.")  +
  theme(legend.position = "none") +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2")
```

\newpage
\subsection{Figure 3: Reactome Pathway Enrichment Analysis}
Enrichment defined from the group of genes <5th percentile of each sigma, or >95th percentile of conservation score.
```{r Fig3, echo=FALSE, fig.height=3.5, comment=NA}
cat("Number of enrichment pathways (sigmas):")
c(sigmaP = length(enrich_sigmaP$Description),
  sigmaPT = length(enrich_sigmaPT$Description),
  sigmaT = length(enrich_sigmaT$Description)) %>%
  sort()

list(
  sigmaP = enrich_sigmaP$Description,
  sigmaPT = enrich_sigmaPT$Description,
  sigmaT = enrich_sigmaT$Description
) %>%
  UpSetR::fromList() %>%
  UpSetR::upset(sets.bar.color = RColorBrewer::brewer.pal(3, "Set2")[c(3, 2, 1)])
grid::grid.text("Figure 3A: Frequency of overlap between pathways that are called \nas most conserved for each hierarchical variance.", x=0.65, y=0.05, gp=grid::gpar(fontsize=10))

cat("Number of enrichment pathways (probs):")
c(score_median = length(enrich_score_median$Description),
  score_median2 = length(enrich_score_median2$Description),
  score_reg = length(enrich_score_reg$Description)) %>%
  sort()

list(
  score_median = enrich_score_median$Description,
  score_median2 = enrich_score_median2$Description,
  score_reg = enrich_score_reg$Description
) %>%
  UpSetR::fromList() %>%
  UpSetR::upset(sets.bar.color = RColorBrewer::brewer.pal(3, "Set2")[c( 2, 1)])
grid::grid.text("Figure 3B: Frequency of overlap between pathways that are called \nas most conserved for each conservation score.", x=0.65, y=0.05, gp=grid::gpar(fontsize=10))
```

\newpage
\subsection{Figure 4: Pathway Enrichment Network}
```{r Fig4, echo=FALSE, message=FALSE, fig.height=2.5}
# Figure 4
cat("sigmaP:")
ggg <- enrich_sigmaP %>%
  ReactomePA::emapplot(showCategory = 30, layout = "graphopt") +
  labs(caption = "Figure 4A: Relationship between pathways that are called as conserved based on sigmaP.", title = NULL)

ggg$layers <- ggg$layers[1:2]

ggg + coord_flip() + 
  ggraph::geom_node_label(aes_(label = ~name), repel = TRUE, size = 2) +
  scale_color_gradientn(colors = "black") +
  guides(color = "none", size = "none")

#ggsave(filename = here("figures", "figure-4.png"), width = 7, dpi = 320, height = 4.326)

# sigmaPT
cat("sigmaPT:")
ggg <- enrich_sigmaPT %>%
  ReactomePA::emapplot(showCategory = 30, layout = "graphopt") +
  labs(caption = "Figure 4B: Relationship between pathways that are called as conserved based on sigmaPT.", title = NULL)

ggg$layers <- ggg$layers[1:2]
ggg + coord_flip() + 
  ggraph::geom_node_label(aes_(label = ~name), repel = TRUE, size = 2) +
  scale_color_gradientn(colors = "black") +
  guides(color = "none", size = "none")

# sigmaT
cat("sigmaT:")
ggg <- enrich_sigmaT %>%
  ReactomePA::emapplot(showCategory = 30, layout = "graphopt") +
  labs(caption = "Figure 4C: Relationship between pathways that are called as conserved based on sigmaT.", title = NULL)

ggg$layers <- ggg$layers[1:2]
ggg + coord_flip() + 
  ggraph::geom_node_label(aes_(label = ~name), repel = TRUE, size = 2) +
  scale_color_gradientn(colors = "black") +
  guides(color = "none", size = "none")
```

\newpage
```{r Fig4DEF, echo=FALSE, message=FALSE, fig.height=2.5}
# score_median
cat("score_median")
ggg <- enrich_score_median %>%
  ReactomePA::emapplot(showCategory = 30, layout = "graphopt") +
  labs(caption = "Figure 4D: Relationship between pathways that are called as conserved based on score_median.", title = NULL)
ggg$layers <- ggg$layers[1:2]
ggg + coord_flip() + 
  ggraph::geom_node_label(aes_(label = ~name), repel = TRUE, size = 2) +
  scale_color_gradientn(colors = "black") +
  guides(color = "none", size = "none")

# score_median2
cat("score_median2")
ggg <- enrich_score_median2 %>%
  ReactomePA::emapplot(showCategory = 30, layout = "graphopt") +
  labs(caption = "Figure 4E: Relationship between pathways that are called as conserved based on score_median2.", title = NULL)
ggg$layers <- ggg$layers[1:2]
ggg + coord_flip() + 
  ggraph::geom_node_label(aes_(label = ~name), repel = TRUE, size = 2) +
  scale_color_gradientn(colors = "black") +
  guides(color = "none", size = "none")

# score_reg
cat("score_reg")
#ggg <- enrich_score_reg %>%
#  ReactomePA::emapplot(showCategory = 30, layout = "graphopt") +
#  labs(caption = "Figure 4G: Relationship between pathways that are called as conserved based on score_reg.", title = NULL)
#ggg$layers <- ggg$layers[1:2]
#ggg + coord_flip() +
#  ggraph::geom_node_label(aes_(label = ~name), repel = TRUE, size = 2) +
#  scale_color_gradientn(colors = "black") +
#  guides(color = "none", size = "none")
```

\subsection{Figure 5: Gene Conservation vs Gene Expression}
```{r Fig5, echo=FALSE, message=FALSE}
# Figure 5A
figure5_data <- permed_data %>% 
  left_join(by = c("UCSC_RefGene_Name" = "gene"), promoted_data) %>%
  left_join(expression_atlas %>% select(`Gene Name`, colon),
            by = c("UCSC_RefGene_Name" = "Gene Name"))

bind_rows(
  figure5_data %>%
    select(sigmaP_v3, sigmaP_pro, colon) %>%
    pivot_longer(-colon) %>%
    drop_na() %>%
    rename(expression = colon),
  figure5_data %>%
    select(sigmaPT_v3, sigmaPT_pro, colon) %>%
    pivot_longer(-colon) %>%
    drop_na() %>%
    rename(expression = colon),
  figure5_data %>%
    select(sigmaT_v3, sigmaT_pro, colon) %>%
    pivot_longer(-colon) %>%
    drop_na() %>%
    rename(expression = colon)
) %>%
  group_by(name) %>%
  mutate(value = factor(11-as.numeric(cut_number(value, 10)))) %>%
  separate(name, c("type", "method"), sep = "_", extra = "merge") %>%
  mutate(type = factor(type, c("sigmaP","sigmaPT","sigmaT"), c("SigmaP-Colon","SigmaPT-Colon","SigmaT-colon")),
         method = factor(method, c("v3", "pro"), c("Boot-strapped values", "Promoter region"))) %>%
  ggplot(aes(value, expression, fill = type)) +
  geom_boxplot() +
  scale_y_log10() +
  guides(color = "none") +
  facet_grid(method ~ type) +
  labs(x = "quantile of conservation by gene", y = "Expression", title = NULL,
       caption = "Figure 5A: Relationship between hierarchical variance (Sigmas) and gene expression.") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2") +
  theme(legend.position = "none")

# Figure 5B
bind_rows(
  figure5_data %>%
    select(score_median_v3, score_median_pro, colon) %>%
    rename(scoremedian_v3 = score_median_v3, scoremedian_pro = score_median_pro) %>%
    pivot_longer(-colon) %>%
    drop_na() %>%
    rename(expression = colon),
  figure5_data %>%
    select(score_median2_v3, score_median2_pro, colon) %>%
    rename(scoremedian2_v3 = score_median2_v3, scoremedian2_pro = score_median2_pro) %>%
    pivot_longer(-colon) %>%
    drop_na() %>%
    rename(expression = colon),
  figure5_data %>%
    select(score_reg_v3, score_reg_pro, colon) %>%
    rename(scorereg_v3 = score_reg_v3, scorereg_pro = score_reg_pro) %>%
    pivot_longer(-colon) %>%
    drop_na() %>%
    rename(expression = colon)
) %>%
  group_by(name) %>%
  mutate(value = factor(as.numeric(cut_number(value, 10)))) %>%
  separate(name, c("type", "method"), sep = "_", extra = "merge") %>%
  mutate(type = factor(type, c("scoremedian","scoremedian2","scorereg"), c("ScoreMedian-colon","ScoreMedian2-colon","Scorereg-colon")),
         method = factor(method, c("v3", "pro"), c("Boot-strapped values", "Promoter region"))) %>%
  ggplot(aes(value, expression, fill = type)) +
  geom_boxplot() +
  scale_y_log10() +
  guides(color = "none") +
  facet_grid(method ~ type) +
  labs(x = "quantile of conservation by gene", y = "Expression", title = NULL,
       caption = "Figure 5B: Relationship between conservation scores and gene expression.") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2") +
  theme(legend.position = "none")
```

<!-- \newpage -->
<!-- \section{Patient-specific analysis} -->
<!-- Basic overview: using a specified number of genes (here 100), take top genes (here using score_prob bootstrapping percentiles) to select a "gene signature" of sorts. Then for each patient, calculate the mean pair-wise distance within the individual patients' tumor samples. Method 1 is these raw values, method 2 normalizes by the mean PWD of all normal samples. Method 3 just examines the overall mean PWD for all tumor samples, normalized by mean PWD of all normal. -->

<!-- ```{r PatientPWD, echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.height=9, fig.width=7} -->
<!-- library(gtools) -->

<!-- # pull out all samples beta values -->
<!-- all_beta <- FullAnnotation[,10:57] -->

<!-- # extract patient and tissue labels -->
<!-- patientLabel <- substr(colnames(all_beta), 1, 1) -->
<!-- patientLabel[10:12] <- "K*" -->
<!-- tissueLabel <- substr(colnames(all_beta), 2, 2) -->
<!-- tissueLabel[!(tissueLabel %in% "N")] <- "T"   #replace A/B/D/M with T -->

<!-- pats = patientLabel %>% sort %>% unique -->
<!-- npat = length(pats) -->

<!-- enrichedGenes <- permed_data %>% filter(score_prob_v3 < quantile(score_prob_v3, 0.05, na.rm = TRUE)) %>% -->
<!--   select(UCSC_RefGene_Name, score_prob, score_prob_v3, n) %>% -->
<!--   rename(Gene=UCSC_RefGene_Name, Pctile=score_prob_v3) -->

<!-- #ngene = 70 # using top N genes -->
<!-- #genes = enrichedGenes %>% top_n(-ngene,Pctile) %>% pull(Gene) -->
<!-- #genes = genes[1:ngene] -->

<!-- genes = enrichedGenes$Gene -->
<!-- ngene = length(genes) -->

<!-- pwd <- function(x) { -->
<!--   mean(dist(x)) -->
<!-- } -->
<!-- pwd2 <- function(x) { -->
<!--   abs(diff(x)) -->
<!-- } -->

<!-- signature_mat1 = matrix(,ngene,npat) # raw tumor pwd -->
<!-- signature_mat2 = matrix(,ngene,npat) # tumor pwd normalized by all normal pwd -->
<!-- for (igene in 1:ngene) { -->
<!--   # each gene, get gene inds, calculate PWD for the tumor samples -->
<!--   geneind = grep(genes[igene], FullAnnotation$UCSC_RefGene_Name) -->

<!--   # calculate PWD of each tumor -->
<!--   tum_pwd <- numeric(npat) -->
<!--   for (ipat in 1:npat) { -->
<!--     # each patient, get tumor sample inds -->
<!--     patind <- which(patientLabel == pats[ipat] & tissueLabel == "T") -->
<!--     tum_pwd[ipat] <- all_beta[geneind,patind] %>% logit %>% apply(1,pwd2) %>% mean -->
<!--   } -->

<!--   # calculate PWD of all normal -->
<!--   norm_pwd <- all_beta[geneind,tissueLabel == "N"] %>% logit %>% apply(1, pwd) %>% mean -->

<!--   signature_mat1[igene,] = tum_pwd -->
<!--   signature_mat2[igene,] = tum_pwd/norm_pwd -->
<!-- } -->

<!-- #cat("  \n\\newline  \nMethod 1 - raw intratumor PWD within each patient") -->
<!-- #signature_mat1 %>% as_tibble %>% `colnames<-`(pats) %>% gather %>% -->
<!-- #  mutate(gene=rep(genes,npat)) %>% rename(Patient=key,PWD=value) %>% -->
<!-- #  ggplot + geom_tile(aes(y=gene,x=Patient,fill=PWD)) + ylab(paste0(ngene," genes")) + -->
<!-- #  theme(axis.text.y = element_blank()) -->
<!-- #cat("  \n\\newline  \nMethod 2 - intratumor PWD normalized by all normal PWD") -->
<!-- #signature_mat2 %>% as_tibble %>% `colnames<-`(pats) %>% gather %>% -->
<!-- #  mutate(gene=rep(genes,npat)) %>% rename(Patient=key,PWD=value) %>% -->
<!-- #  ggplot + geom_tile(aes(y=gene,x=Patient,fill=PWD)) + ylab(paste0(ngene," genes")) + -->
<!-- #  theme(axis.text.y = element_blank()) -->
<!-- #cat("  \n\\newline  \nMethod 3 - all tumor PWD normalized by all normal PWD") -->
<!-- #signature_mat3 %>% as_tibble %>% mutate(gene=genes) %>% rename(PWD=V1) %>% -->
<!-- #  ggplot + geom_tile(aes(y=gene,x=1,fill=PWD)) + ylab(paste0(ngene," genes")) + -->
<!-- #  theme(axis.text.y = element_blank()) -->

<!-- library(pheatmap) -->
<!-- library(dendsort) -->

<!-- cat("  \n\\newline  \nMethod 1 - raw intratumor PWD within each patient") -->
<!-- df1 = as.data.frame(signature_mat1) -->
<!-- rownames(df1) <- genes -->
<!-- colnames(df1) <- pats -->
<!-- sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...))) -->
<!-- mat_cluster_cols <- sort_hclust(hclust(dist(t(df1)))) -->
<!-- mat_cluster_rows <- sort_hclust(hclust(dist(df1))) -->
<!-- pheatmap( -->
<!--   mat=df1, -->
<!--   cluster_cols = mat_cluster_cols, -->
<!--   cluster_rows = mat_cluster_rows, -->
<!--   show_rownames = FALSE, -->
<!--   show_colnames = TRUE, -->
<!--   main = "Top 1118 Genes: Tumor PWD" -->
<!-- ) -->

<!-- cat("  \n\\newline  \nMethod 2 - intratumor PWD normalized by all normal PWD") -->
<!-- df1 = as.data.frame(signature_mat2) -->
<!-- rownames(df1) <- genes -->
<!-- colnames(df1) <- pats -->
<!-- sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...))) -->
<!-- mat_cluster_cols <- sort_hclust(hclust(dist(t(df1)))) -->
<!-- mat_cluster_rows <- sort_hclust(hclust(dist(df1))) -->
<!-- pheatmap( -->
<!--   mat=df1, -->
<!--   cluster_cols = mat_cluster_cols, -->
<!--   cluster_rows = mat_cluster_rows, -->
<!--   show_rownames = FALSE, -->
<!--   show_colnames = TRUE, -->
<!--   main = "Top 1118 Genes: Tumor PWD" -->
<!-- ) -->
<!-- ``` -->

\newpage
\section{Appendix: Enrichment Gene/Pathway Lists}
\subsection{A1: SigmaP}
Genes are selected with cutoff <5th percentile (adjusted bootstrap), showing top 30 genes/pathways
```{r App_sigmaP, echo=FALSE, fig.fullwidth=TRUE, comment=NA, message=FALSE}
enrichedGenes <- permed_data %>%
  select(UCSC_RefGene_Name, sigmaP, sigmaP_v3, n) %>%
  rename(Gene=UCSC_RefGene_Name, Pctile=sigmaP_v3) %>%
  filter(Pctile < quantile(Pctile, 0.05, na.rm = TRUE)) %>%
  arrange(Pctile)

enrichedPaths <- tibble(Pathway = enrich_sigmaP$Description, ngenes = enrich_sigmaP$Count, padj = enrich_sigmaP$p.adjust) %>% arrange(padj)

ngene <- dim(enrichedGenes)[1]
npath <- length(enrich_sigmaP$Description)

cat(paste0("Total number of genes: ", ngene, "     Number of enrichment pathways: ", npath))

nshow = 30
if (ngene > nshow) {
  enrichedGenes <- enrichedGenes[1:nshow,]
}
if (npath > nshow) {
  enrichedPaths <- enrichedPaths[1:nshow,]
}

library(pander)
enrichedGenes %>% pander()

cat("SigmaP top 30 enriched paths:  \n")
enrichedPaths %>% pander()
```

\newpage
\subsection{A2: SigmaPT}
Genes are selected with cutoff <5th percentile (adjusted bootstrap), showing top 30 genes/pathways
```{r App_sigmaPT, echo=FALSE, comment=NA, message=FALSE}
enrichedGenes <- permed_data %>%
  select(UCSC_RefGene_Name, sigmaPT, sigmaPT_v3, n) %>%
  rename(Gene=UCSC_RefGene_Name, Pctile=sigmaPT_v3) %>%
  filter(Pctile < quantile(Pctile, 0.05, na.rm = TRUE)) %>%
  arrange(Pctile)

enrichedPaths <- tibble(Pathway = enrich_sigmaPT$Description, ngenes = enrich_sigmaPT$Count, padj = enrich_sigmaPT$p.adjust) %>% arrange(padj)

ngene <- dim(enrichedGenes)[1]
npath <- length(enrich_sigmaPT$Description)

cat(paste0("Total number of genes: ", ngene, "     Number of enrichment pathways: ", npath))

nshow = 30
if (ngene > nshow) {
  enrichedGenes <- enrichedGenes[1:nshow,]
}
if (npath > nshow) {
  enrichedPaths <- enrichedPaths[1:nshow,]
}

enrichedGenes %>% pander()

cat("\nSigmaPT top 30 enriched paths:  \n")
enrichedPaths %>% pander()
```

\newpage
\subsection{A3: SigmaT}
Genes are selected with cutoff <5th percentile (adjusted bootstrap), showing top 30 genes/pathways
```{r App_sigmaT, echo=FALSE, comment=NA, message=FALSE}
enrichedGenes <- permed_data %>%
  select(UCSC_RefGene_Name, sigmaT, sigmaT_v3, n) %>%
  rename(Gene=UCSC_RefGene_Name, Pctile=sigmaT_v3) %>%
  filter(Pctile < quantile(Pctile, 0.05, na.rm = TRUE)) %>%
  arrange(Pctile)

enrichedPaths <- tibble(Pathway = enrich_sigmaT$Description, ngenes = enrich_sigmaT$Count, padj = enrich_sigmaT$p.adjust) %>% arrange(padj)

ngene <- dim(enrichedGenes)[1]
npath <- length(enrich_sigmaT$Description)

cat(paste0("Total number of genes: ", ngene, "     Number of enrichment pathways: ", npath))

nshow = 30
if (ngene > nshow) {
  enrichedGenes <- enrichedGenes[1:nshow,]
}
if (npath > nshow) {
  enrichedPaths <- enrichedPaths[1:nshow,]
}

enrichedGenes %>% pander()

cat("SigmaT top 30 enriched paths:  \n")
enrichedPaths %>% pander()
```

<!-- \newpage -->
<!-- \subsection{A4: Score Prob} -->
<!-- Genes are selected with cutoff >95th percentile (adjusted bootstrap), showing top 50 genes/pathways -->
<!-- ```{r App_score_prob, echo=FALSE, comment=NA, message=FALSE} -->
<!-- enrichedGenes <- permed_data %>%  -->
<!--   select(UCSC_RefGene_Name, score_prob, score_prob_v3, n) %>% -->
<!--   rename(Gene=UCSC_RefGene_Name, Score_median=score_prob, Pctile=score_prob_v3) %>% -->
<!--   filter(Pctile > quantile(Pctile, 0.95, na.rm = TRUE)) %>% -->
<!--   arrange(Pctile) -->

<!-- enrichedPaths <- tibble(Pathway = enrich_score_prob$Description, ngenes = enrich_score_prob$Count, padj = enrich_score_prob$p.adjust) %>% arrange(padj) -->

<!-- ngene <- dim(enrichedGenes)[1] -->
<!-- npath <- length(enrich_score_prob$Description) -->

<!-- cat(paste0("Total number of genes: ", ngene, "     Number of enrichment pathways: ", npath)) -->

<!-- nshow = 50 -->
<!-- if (ngene > nshow) { -->
<!--   enrichedGenes <- enrichedGenes[1:nshow,] -->
<!-- } -->
<!-- if (npath > nshow) { -->
<!--   enrichedPaths <- enrichedPaths[1:nshow,] -->
<!-- } -->

<!-- enrichedGenes %>% pander() -->

<!-- cat("Score_prob top 50 enriched paths:  \n") -->
<!-- enrichedPaths %>% pander() -->
<!-- ``` -->

<!-- \newpage -->
<!-- \subsection{A5: Score Mean} -->
<!-- Genes are selected with cutoff >95th percentile (adjusted bootstrap), showing top 50 genes/pathways -->
<!-- ```{r App_score_mean, echo=FALSE, comment=NA, message=FALSE} -->
<!-- enrichedGenes <- permed_data %>%  -->
<!--   select(UCSC_RefGene_Name, score_mean, score_mean_v3, n) %>% -->
<!--   rename(Gene=UCSC_RefGene_Name, Score_median=score_mean, Pctile=score_mean_v3) %>% -->
<!--   filter(Pctile > quantile(Pctile, 0.95, na.rm = TRUE)) %>% -->
<!--   arrange(Pctile) -->

<!-- enrichedPaths <- tibble(Pathway = enrich_score_mean$Description, ngenes = enrich_score_mean$Count, padj = enrich_score_mean$p.adjust) %>% arrange(padj) -->

<!-- ngene <- dim(enrichedGenes)[1] -->
<!-- npath <- length(enrich_score_mean$Description) -->

<!-- cat(paste0("Total number of genes: ", ngene, "     Number of enrichment pathways: ", npath)) -->

<!-- nshow = 50 -->
<!-- if (ngene > nshow) { -->
<!--   enrichedGenes <- enrichedGenes[1:nshow,] -->
<!-- } -->
<!-- if (npath > nshow) { -->
<!--   enrichedPaths <- enrichedPaths[1:nshow,] -->
<!-- } -->

<!-- enrichedGenes %>% pander() -->

<!-- cat("Score_mean top 50 enriched paths:  \n") -->
<!-- enrichedPaths %>% pander() -->
<!-- ``` -->

\newpage
\subsection{A6: Score Median}
Genes are selected with cutoff >95th percentile (adjusted bootstrap), showing top 50 genes/pathways
```{r App_score_median, echo=FALSE, comment=NA, message=FALSE}
enrichedGenes <- permed_data %>% 
  select(UCSC_RefGene_Name, score_median, score_median_v3, n) %>%
  rename(Gene=UCSC_RefGene_Name, Score_median=score_median, Pctile=score_median_v3) %>%
  filter(Pctile > quantile(Pctile, 0.95, na.rm = TRUE)) %>%
  arrange(Pctile)

enrichedPaths <- tibble(Pathway = enrich_score_median$Description, ngenes = enrich_score_median$Count, padj = enrich_score_median$p.adjust) %>% arrange(padj)

ngene <- dim(enrichedGenes)[1]
npath <- length(enrich_score_median$Description)

cat(paste0("Total number of genes: ", ngene, "     Number of enrichment pathways: ", npath))

nshow = 50
if (ngene > nshow) {
  enrichedGenes <- enrichedGenes[1:nshow,]
}
if (npath > nshow) {
  enrichedPaths <- enrichedPaths[1:nshow,]
}

enrichedGenes %>% pander()

cat("Score_median top 50 enriched paths:  \n")
enrichedPaths %>% pander()
```

\newpage
\subsection{A7: Score Median2}
Genes are selected with cutoff >95th percentile (adjusted bootstrap), showing top 50 genes/pathways
```{r App_score_median2, echo=FALSE, comment=NA, message=FALSE}
enrichedGenes <- permed_data %>% 
  select(UCSC_RefGene_Name, score_median2, score_median2_v3, n) %>%
  rename(Gene=UCSC_RefGene_Name, Score_median2=score_median2, Pctile=score_median2_v3) %>%
  filter(Pctile > quantile(Pctile, 0.95, na.rm = TRUE)) %>%
  arrange(Pctile)

enrichedPaths <- tibble(Pathway = enrich_score_median2$Description, ngenes = enrich_score_median2$Count, padj = enrich_score_median2$p.adjust) %>% arrange(padj)

ngene <- dim(enrichedGenes)[1]
npath <- length(enrich_score_median2$Description)

cat(paste0("Total number of genes: ", ngene, "     Number of enrichment pathways: ", npath))

nshow = 50
if (ngene > nshow) {
  enrichedGenes <- enrichedGenes[1:nshow,]
}
if (npath > nshow) {
  enrichedPaths <- enrichedPaths[1:nshow,]
}

enrichedGenes %>% pander()

cat("Score_median2 top 50 enriched paths:  \n")
enrichedPaths %>% pander()
```

\newpage
\subsection{A9: Score Regularized}
Genes are selected with cutoff >95th percentile (adjusted bootstrap), showing top 50 genes/pathways
```{r App_score_reg, echo=FALSE, comment=NA, message=FALSE}
enrichedGenes <- permed_data %>% 
  select(UCSC_RefGene_Name, score_reg, score_reg_v3, n) %>%
  rename(Gene=UCSC_RefGene_Name, Score_reg=score_reg, Pctile=score_reg_v3) %>%
  filter(Pctile > quantile(Pctile, 0.95, na.rm = TRUE)) %>%
  arrange(Pctile)

enrichedPaths <- tibble(Pathway = enrich_score_reg$Description, ngenes = enrich_score_reg$Count, padj = enrich_score_reg$p.adjust) %>% arrange(padj)

ngene <- dim(enrichedGenes)[1]
npath <- length(enrich_score_reg$Description)

cat(paste0("Total number of genes: ", ngene, "     Number of enrichment pathways: ", npath))

nshow = 50
if (ngene > nshow) {
  enrichedGenes <- enrichedGenes[1:nshow,]
}
if (npath > nshow) {
  enrichedPaths <- enrichedPaths[1:nshow,]
}

enrichedGenes %>% pander()

cat("Score_reg top 50 enriched paths:  \n")
enrichedPaths %>% pander()
```