### 2020-10-01 prior_sensitivity.Rmd

---
title: "DNAm Hierarchical Model - Prior Sensitivity Analysis "
author: "Kevin Murgas"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: true
    number_section: true
    toc: true
---

\newpage
Prior sensitivity is assessed by looking at Stan fit results of different sets of priors, each with an increasing level of variance relaxation. Current data: 1x (original), 2x, 3x, and 5x.
\newline Priors are assessed by 3 metrics: n_eff and Rhat to assess Stan MCMC convergence, and Kullback-Leibler divergence to assess prior constraints.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(gtools)

# set working directory
setwd("/Users/kevinmurgas/Documents/Data+ project/EPIC data")
library(here)

select <- dplyr::select
rename <- dplyr::rename
filter <- dplyr::filter
slice <- dplyr::slice

set.seed(1234)
```

```{r prior_sensitivity, echo=FALSE}

for (i in c(1:4)) {
  stanFile = "FullResultsTCGA_relaxgamma3.Rdata"
  stanFile = switch(i,
                    "FullResultsTCGA_gamma.Rdata",
                    "FullResultsTCGA_relaxgamma2.Rdata",
                    "FullResultsTCGA_relaxgamma3.Rdata",
                    "FullResultsTCGA_relaxgamma5.Rdata")
  print(stanFile)
  #load(here(stanFile))
}

```