### 2020-10-20 differential_methylation.Rmd

---
title: "DNAm Hierarchical Model - Fixed Effects Differential Methylation"
author: "Kevin Murgas"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: true
    number_section: true
    toc: true
---

\newpage
\section{Differential Methylation}
Differential methylation is typically defined using a basic linear model over the data points between normal and tumor samples.
$$ M_i = \mu + \Delta M*\text{tInd}(i) $$
Where tInd is a binary indicator for tumor samples.
\newline This linear fit captures the normal mean (intercept) and shift in methylation in tumors (slope). The slope ($\Delta M$) is used to define differentially methylated regions by first z-score normalizing and then determining statistical significance using a permutation based false discovery rate (FDR)
$$ z = \text{mean}(\Delta M) / \text{s.d.}(\Delta M) $$
\newline I am currently a bit unclear on the FDR step. Irizarry et al 2009 reference : Efron B, Tibshirani R, Storey JD, Tusher V. Empirical Bayes analysis of a microarray experiment. Journal of the American statistical association. 2001 Dec 1;96(456):1151-60.
\newline Z-scores can convert to p-values, and Benjamini-Hochberg step-up can account for multiple hypothesis correction. The permutation could be from shuffling the normal/tumor labels on 1000 draws.

\newline The minfi dmpFinder approach uses a linear fit (limma package), and computes a t-statistic for the slope by the following formula:
$$ t = (\Delta M) / (\sigma_{\Delta M} * \sigma_{resid}) $$
Which is then converted to a p-value using the t-distribution with df= # samples - 2. The p-value is then corrected by Storey et al. 2003 Bayesian q-value method, which is similar to Benjamini-Hochberg linear step-up FDR procedure but divides by a factor of an estimated prior.

\section{Model Description: Fixed Effects}
\subsection{Model equation}
Let i=sample, j=patient
$$ y_i = \mu + a_j +  \text{tInd}(i)*(\beta_T + b_j + c_i) $$
$$ a_j \sim N(0,\sigma_P) , b_j \sim N(0,\sigma_{PT}) , c_i \sim N(0,\sigma_T) $$
\subsection{Differential Methylation using betaT}
Effects $\mu$ and $\beta_T$ describe the normal mean methylation and tumor shift in methylation, respectively. Therefore, differentially methylated CpG sites will have significant betaT values. A similar procedure is used as the $\Delta M$ , using a z-score normalization, and permutation based FDR.
\newline This provides a way to identify CpG sites with significant differential methylation. These sites can then be directly compared to C-DMRs of Irizarry 2009 (provided as regions, i.e. base ranges on a given chromosome), or go up to identify genes with either any d.m. CpG sites or a certain fraction. Genes can also be compared to Irizarry C-DMRs (each region has a single gene listed)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(gtools)

# set working directory
setwd("/Users/kevinmurgas/Documents/Data+ project/EPIC data")
library(here)

select <- dplyr::select
rename <- dplyr::rename
filter <- dplyr::filter
slice <- dplyr::slice

set.seed(1234)
```

\newpage
\section{Compute DM - raw beta values}
```{r dm_beta, echo=FALSE}
# Load raw methylation data
load(here("myFA_bulkonly.Rdata"))

# extract beta values from data
all_beta <- FullAnnotation[,10:57]
rownames(all_beta)<-FullAnnotation$IlmnID
tissueLabel <- substr(colnames(all_beta), 2, 2)
tissueLabel[!(tissueLabel %in% "N")] <- "T"

# convert to M values, tInd
all_M <- logit(all_beta) # base e
#all_M <- log2(all_beta/(1-all_beta)) # base 2
tInd <- tissueLabel=="T"

nsamp = dim(all_beta)[2]
ymean = rowSums(all_M)/nsamp
SStot = rowSums( (all_M - ymean)^2 )

# same procedure as minfi dmpFinder
# use limma to compute linear fits
library(limma)
design <- cbind(muM=1,deltaM=tInd)
fit <- lmFit(all_M, design)
tstat.ord <- fit$coef[,2] / fit$stdev.unscaled[,2] / fit$sigma
pval <- 2*pt(abs(tstat.ord), fit$df.residual, lower.tail=FALSE)
qvalBH <- p.adjust(pval, method="BH")

# qvalue approach used by minfi dmpFinder (from Storey 2003)
library(siggenes)
p0 <- pi0.est(pval[!is.na(pval)])$p0
qvalS <- qvalue.cal(pval, p0)

fit_stats <- cbind(as.data.frame(fit), tstat.ord, pval, qvalBH, p0, qvalS)

## using emprical Bayes stats (not necessary unless <10 samples)
#stats <- as.data.frame(eBayes(fit)) %>%
#  select(-c(stdev.unscaled.muM, stdev.unscaled.deltaM)) %>% #df.residual, df.total 
#  mutate(t.ord.muM=tstat.ord[,1], t.ord.deltaM=tstat.ord[,2]) %>%
#  mutate(qval = p.adjust(pt.ord.deltaM, method="BH"))
## for some reason the qval is off from dmpfinder qval by a constant factor 3.496509

## minfi method
# library(minfi)
# dmpsM <- dmpFinder(as.matrix(all_M), tInd, type="continuous")
# dmpsM_new <- dmpsM[ order(row.names(dmpsM)), ]

library(gt)
data.frame(sumBH = sum(qvalBH<0.05), sumS = sum(qvalS<0.05)) %>%
  rename('# Signif. sites (Benjamini-Hochberg)' = sumBH, '# Signif. sites (Storey)' = sumS)
```

\section{Compute DM - Stan fit betaT}
Use a similar metric, with a t-score based on the median estimate and SEM of posterior. Unclear on how to run significance bootstrapping, but should be find with using the t-test p-value and FDR correction with B-H step up. 
\newline Second approach would be use a z-score based on all of the estimates.
```{r dm_stan, echo=FALSE}
# load Stan fits
stanFile = "FullResultsTCGA_relaxgamma3.Rdata"
load(here(stanFile))

nsamp = dim(all_beta)[2]
sd_all = sd(betaT_full$p50)
betaT_stats <- betaT_full %>% mutate(tstat = p50/(sem*sqrt(n_eff))) %>%
#  mutate(SSreg = sem*n_eff) %>%
#  mutate(SSres = SStot - SSreg) %>%
  mutate(pval = 2*pt(abs(tstat), nsamp-2, lower.tail=FALSE)) %>%
  mutate(qvalBH = p.adjust(pval, method="BH")) %>%
  mutate(zscore = p50/sd_all) %>%
  mutate(pz = 2*pnorm(abs(zscore),lower.tail=FALSE)) %>%
  mutate(qvalBHz = p.adjust(pz, method="BH"))
```

\section{Compare to Irizarry et al 2009}
Irizarry use a similar differential methylation profiling approach in colorectal cancers (different methylation assay), and identify numerous cancer-specific diff. meth. regions (C-DMRs). Compare their list, which reports ranges of bases with single gene labels, to our differential methylation findings.
\newline Doi et al. 2009 (same group as Irizarry 2009), perform a similar analysis to determine C-DMRs in colorectal cancer. These C-DMRs are reported in the "DMR" column of the Illumina EPIC array manifest.
\newline Summarize % of overlap in each case, using specific bases of CpG sites or using gene names.
```{r compare_irizarry, echo=FALSE}
DMR_data = (fit_stats$qvalBH < 0.05) & (abs(fit_stats$coefficients.deltaM) > 1.4)
DMR_beta = (betaT_stats$qvalBH < 0.05) & (abs(betaT_stats$p50) > 1.4)
CDMR_epic = (FullAnnotation$DMR == "CDMR") #| (FullAnnotation$DMR == "DMR")

cor(fit_stats$tstat.ord,betaT_stats$tstat)
cor(fit_stats$coefficients.deltaM,betaT_stats$p50)

# DMR_list = cbind(DMR_data, CDMR_epic)
# crosstab <- xtabs(~DMR_data + CDMR_epic, DMR_list)
# crosstab

# DMR_list = cbind(DMR_beta, CDMR_epic)
# crosstab <- xtabs(~DMR_beta + CDMR_epic, DMR_list)
# crosstab

DMR_list = cbind(DMR_beta, DMR_data)
crosstab <- xtabs(~DMR_beta + DMR_data, DMR_list)
crosstab

chisq.test(crosstab)
fisher.test(crosstab)
OR = (crosstab[1,1]+crosstab[2,2])/(crosstab[1,2]+crosstab[2,1])
phi = (prod(c(crosstab[1,1],crosstab[2,2]))-prod(crosstab[1,2],crosstab[2,1]))/sqrt(prod(rowSums(crosstab))*prod(rowSums(t(crosstab))))

# a few correlations between linear and betaT methods
cor(fit_stats$tstat.ord,betaT_stats$tstat)
cor(fit_stats$coefficients.deltaM,betaT_stats$p50)

```