### 2020-11-24 Paper_Figures.Rmd
Edits to figures

---
title: "DNA Methylation Hierarchical Variance Model\nPaper Figures"
author: "Kevin Murgas"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: true
    number_section: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)

library(gt)
library(rlang)
library(knitr)
library(gtools)

# set working directory
setwd("/Users/kevinmurgas/Documents/Data+ project/EPIC data")
library(here)

select <- dplyr::select
rename <- dplyr::rename
filter <- dplyr::filter
slice <- dplyr::slice

set.seed(1234)
```

```{r Load_data, echo=FALSE, include=TRUE, message=FALSE}
# Load raw methylation data
load(here("myFA_bulkonly.Rdata"))

stanFile = "FullResultsTCGA_relaxgamma3.Rdata"
load(here(stanFile))

prepdatafolder = "data_final_rg3"
cat(sprintf("Load prepped data dir: %s (needs to be pre-processed)...",prepdir))
gened_data <- read_csv(here(prepdatafolder, "gened_data.csv"), col_types = list("CHR" = col_character()))
prepped_data <- read_csv(here(prepdatafolder, "prepped_data.csv"))
permed_data <- read_csv(here(prepdatafolder, "permed_data.csv"))

# # plot beta/M values
# all_beta = FullAnnotation[,-c(1,2,3,4,5,6,7,8,9)]
# mean_beta = rowMeans(all_beta)
# all_M = logit(all_beta)
# mean_M = rowMeans(all_M)
# p_beta <- data.frame(value=mean_beta) %>% gather %>% ggplot() + geom_histogram(aes(x=value),binwidth=0.01,fill="lightgray",color="grey50") + theme_classic() + xlim(c(0,1)) + theme(aspect.ratio = 1) + xlab("Beta value")
# p_M <- data.frame(value=mean_M) %>% gather %>% ggplot() + geom_histogram(aes(x=value),binwidth=0.1,fill="lightgray",color="grey50") + theme_classic() + xlim(c(-5,5)) + theme(aspect.ratio = 1) + xlab("M value")
# p_BM <- ggarrange(p_beta, p_M, nrow = 1, ncol = 2)
# # ggsave("betaMdists.svg",p_BM)
```

\newpage
\section{Model Overview, Priors, Example Site Fit}

\subsection{Data}
Logit-transformed beta values, from EPIC methylation array pre-processed with minfi Noob algorithm. \newline
\newline Sample counts (total): 48 samples from 21 patients. \newline
By tissue: 6 normal samples, 42 paired bulk tumor samples.

\subsection{Model equation}
Let i=patient, j=sample (0=normal, 1,2=tumor sample 1,2)
$$ y_{ij} = \mu + \alpha_i +  (1-\delta_{j0})*(\nu + \beta_i + \gamma_{ij}) + \epsilon_{ij} $$
$$ \alpha_i \sim N(0,\sigma_P) , \beta_i \sim N(0,\sigma_{PT}) , \gamma_{ij} \sim N(0,\sigma_T) $$
\newline Model is fit at each individual CpG site using RStan's MCMC sampling algorithm. \newline
Stan Parameters: 4 chains x 2000 iterations (200 warmup) = 7200 total draws. adapt_delta = 0.999

\subsection{Prior Probabilities}
Priors for each parameter (mu, nu, sigmaP/PT/T) were defined using TCGA 450k array data from colorectal tumor patients with paired multiple tumor and/or normal+tumor samples. \newline Empirical values of mean and standard deviations were assesed for all 450k CpG sites. \newline The resulting distributions were fit with: bimodal gaussian (mu), Cauchy (betaT), or gamma distributions (sigmas). \newline
SigmaE (error variance) prior was defined using a non-informative gamma distribution: gamma(2,2).
\newline Additionally: Prior relaxation was performed by algebraically increasing the variance of the distributions (or each Gaussian component in the case of mu) while maintaining the same mode (peak).

```{r Priors, echo=FALSE, fig.fullwidth=TRUE, warning=FALSE, comment=NA}
# load TCGA estimates for mu, betaT, sigmaP, sigmaPT, sigmaT
load(file="/Users/kevinmurgas/Documents/Data+ project/TCGA data/estimates_allpats.Rdata")

# remove all zeros, and take square root of variance (stored variable) to get sigma/stdev
sigmaP_ests <- sqrt(sigmaP_ests[which(!!sigmaP_ests)])
sigmaPT_ests <- sqrt(sigmaPT_ests[which(!!sigmaPT_ests)])
sigmaT_ests <- sqrt(sigmaT_ests[which(!!sigmaT_ests)])
mu_ests <- mu_ests[which(!!mu_ests)]
betaT_ests <- betaT_ests[which(!!betaT_ests)]

# for each parameter, plot the distribution and overlay the fits
fBG <- function(x, a, m1, s1, m2, s2) a*dnorm(x,m1,s1) + (1-a)*dnorm(x,m2,s2) # bimodal gaussian
post_prior_histogram <- function(vals,priorfun1,priorfun2,bw,xl,titlestr,legFlag=FALSE) {
  # stacked histogram colored by gene-associated sites
  # and prior curve overlaid
  plot_df <- data.frame(val=vals)
  p1 <- ggplot(plot_df, aes(x=val)) +
    geom_histogram(aes(y=..count../(length(vals)*bw), fill="Empirical"), binwidth=bw) +
    stat_function(fun=function(x) priorfun1(x),geom="line",aes(color="Fit")) +
    stat_function(fun=function(x) priorfun2(x),geom="line",aes(color="3x Relax")) +
    ggtitle(titlestr) + xlim(xl) + theme_classic() + xlab(titlestr) + ylab("Freq.") +
    scale_colour_manual(NULL, values = c("Empirical" = "gray","Fit" ="red","3x Relax" = "blue")) +
    scale_fill_manual(NULL,values = c("Empirical"="gray")) +
    theme(legend.position = c(0.75, 0.65))
  if (!legFlag) {p1 <- p1 + theme(legend.position = "none")}
  p1
}

# plot mu,betaT,sigmaE,sigmaP,sigmaPT,sigmaT with priors overlaid
pM <- post_prior_histogram(mu_ests, function(x) fBG(x,0.3557,-3.0675, 0.75214,1.1290, 1.39197), function(x) fBG(x,0.3557,-3.0675, sqrt(3)*0.75214,1.1290, sqrt(3)*1.39197), 0.1, c(-5,5), "Mu")
pB <- post_prior_histogram(betaT_ests, function(x) dcauchy(x,-0.04051, 0.19330), function(x) dcauchy(x,-0.04051, 0.5799), 0.1, c(-5,5), "Nu")
pP <- post_prior_histogram(sigmaP_ests, function(x) dgamma(x, 5.1989, 16.6192), function(x) dgamma(x, 2.7693, 7.0029), 0.05, c(0,2), "SigmaP")
pPT <- post_prior_histogram(sigmaPT_ests, function(x) dgamma(x, 3.5539, 5.5324), function(x) dgamma(x, 2.1457, 2.4819), 0.05, c(0,2), "SigmaPT")
pT <- post_prior_histogram(sigmaT_ests, function(x) dgamma(x, 6.8383, 25.7560), function(x) dgamma(x, 3.3643, 10.430), 0.05, c(0,2), "SigmaT",legFlag=TRUE)
pE <- ggplot() + stat_function(fun=function(x) dgamma(x, 2, 2),geom="line",color="blue") + ggtitle("SigmaE") + xlim(c(0,2)) + theme_classic() + theme(legend.position = "none") + xlab("SigmaE") + ylab("Freq.")

pBlank <- ggplot(data=data.frame(len=c(1,2)),aes(x=len)) + theme_void() + geom_bar(aes(fill="Empirical")) +
    stat_function(fun=function(x) 3,geom="line",aes(color="1x variance")) +
    stat_function(fun=function(x) 4,geom="line",aes(color="3x variance")) + xlim(0,1) + ylim(0,1) + scale_colour_manual(NULL, values = c("Empirical" = "gray","1x variance" ="red","3x variance" = "blue")) + scale_fill_manual(NULL,values = "gray") + theme(legend.position = c(2, 0.75))

p_all <- ggarrange(pM,pB,pE,pP,pPT,pT,hjust=0,nrow=2,ncol=3)
print(p_all)
rm(pM,pB,pBlank,pP,pPT,pT,pE)

```

\subsection{CpG Conservation score}
We want to understand which sites of DNA methylation are fundamentally conserved within tumors, i.e. lower variation in methlyation within tumor (sigmaT) relative to the variation within normal healthy tissue (sigmaP). Therefore we choose the log-ratio (base 2) of sigmaP/sigmaT to represent a relative conservation score, which is positive when tumor variation is lower than normal variation. We take the median of the posterior distribution constructed from the two sigmaP,T posteriors for this score.

$$\text{score}_\text{med} = \text{Median} (\text{log}_2\frac{\sigma_{P}}{\sigma_T}) $$

\newpage
\subsection{Figure 2A: Example site fit, 2B: Posteriors}
![Fit](/Users/kevinmurgas/Documents/Data+ project/Manuscript/example_fit.png)
\newline \newline
![Posteriors](/Users/kevinmurgas/Documents/Data+ project/Manuscript/example_posteriors.png)
\newpage
\section{Figure 2: Model Fit Results}
\subsection{Parameter Distributions}
Histograms are median values of posterior distribution
```{r Params_site, echo=FALSE, fig.fullwidth=TRUE, warning=FALSE, comment=NA}
genemask = !(FullAnnotation$UCSC_RefGene_Name=="")
cat(sprintf("N sites: total=%i, gene-associated=%i, non-gene=%i  \n",length(genemask),sum(genemask),length(genemask)-sum(genemask)))

fBG <- function(x, a, m1, s1, m2, s2) a*dnorm(x,m1,s1) + (1-a)*dnorm(x,m2,s2) # bimodal gaussian
post_prior_histogram <- function(postvals,mask,priorfun,bw,xl,titlestr) {
  # stacked histogram colored by gene-associated sites
  # and prior curve overlaid
  plot_df <- data.frame(val=postvals, gene_assoc=mask)
  p1 <- ggplot(plot_df, aes(x=val)) +
    geom_histogram(aes(y=..count../length(mask),fill=gene_assoc), binwidth=bw) +
    stat_function(fun=function(x) bw*priorfun(x),geom="line",color="black") +
    ggtitle(titlestr) + xlim(xl) + theme_classic() + theme(legend.position = "none") + xlab(titlestr) + ylab("Freq.")
  p1
}

# plot mu,betaT,sigmaE,sigmaP,sigmaPT,sigmaT with priors overlaid
pM <- post_prior_histogram(mu_full$p50, genemask, function(x) 1.5*fBG(x,0.3557,-3.0675, 0.75214,1.1290, 1.39197), 0.1, c(-5,5), "Mu")
pB <- post_prior_histogram(betaT_full$p50, genemask, function(x) 1.7*dcauchy(x,-0.04051, 0.5799), 0.1, c(-5,5), "BetaT")
pE <- post_prior_histogram(sigmaE_full$p50, genemask, function(x) 4*dgamma(x,2,2), 0.05, c(0,2), "SigmaE")
pP <- post_prior_histogram(sigmaP_full$p50, genemask, function(x) 2*dgamma(x, 2.7693, 7.0029), 0.05, c(0,2), "SigmaP")
pPT <- post_prior_histogram(sigmaPT_full$p50, genemask, function(x) 1.8*dgamma(x, 2.1457, 2.4819), 0.05, c(0,2), "SigmaPT")
pT <- post_prior_histogram(sigmaT_full$p50, genemask, function(x) 1.2*dgamma(x, 3.3643, 10.430), 0.05, c(0,2), "SigmaT")
# log sigmas
pLP <- post_prior_histogram(log(sigmaP_full$p50), genemask, function(x) 0.52*dgamma(exp(x), 2.7693, 7.0029), 0.1, c(-5,5), "log(SigmaP)")
pLPT <- post_prior_histogram(log(sigmaPT_full$p50), genemask, function(x) 0.7*dgamma(exp(x), 2.1457, 2.4819), 0.1, c(-5,5), "log(SigmaPT)")
pLT <- post_prior_histogram(log(sigmaT_full$p50), genemask, function(x) 0.4*dgamma(exp(x), 3.3643, 10.430), 0.1, c(-5,5), "log(SigmaT)") + theme(legend.position = c(0.8, 0.6), legend.text=element_text(size=6), legend.title=element_text(size=6))

# arrange each parameter as a subplot # ,labels=c("M","B","E","P","PT","T","lP","lPT","lT")
p_all <- ggarrange(pM,pB,pE,pP,pPT,pT,pLP,pLPT,pLT,hjust=0,nrow=3,ncol=3)
print(p_all)
rm(pM,pB,pE,pP,pPT,pT,pLP,pLPT,pLT)
```

\subsection{Conservation score}
```{r Fig1B, echo=FALSE, fig.height=2, fig.width=6.5, fig.fullwidth=TRUE, warning=FALSE}

post_histogram <- function(postvals,mask,bw,xl,titlestr) {
  # stacked histogram colored by gene-associated sites
  # and prior curve overlaid
  plot_df <- data.frame(val=postvals, gene_assoc=mask)
  p1 <- ggplot(plot_df, aes(x=val)) +
    geom_histogram(aes(y=..count../length(mask),fill=gene_assoc), binwidth=bw) +
    ggtitle(titlestr) + xlim(xl) + theme_classic() + theme(legend.position = "none") + xlab(titlestr) + ylab("Freq.")
}

# overlaid transparent histograms
post_histogram <- function(postvals,mask,bw,titlestr) {
  # stacked histogram colored by gene-associated sites
  # and prior curve overlaid
  plot_df <- data.frame(val=postvals, gene_assoc=mask)
  gene_assoc_df <- plot_df %>% filter(gene_assoc)
  not_gene_assoc_df <- plot_df %>% filter(!gene_assoc)
  p1 <- ggplot(plot_df, aes(x=val)) +
    geom_histogram(data=not_gene_assoc_df, aes(y=..density..,fill="#00BFC4",alpha=0.4), binwidth=bw) +
    geom_histogram(data=gene_assoc_df, aes(y=..density..,fill="#F8766D",alpha=0.4), binwidth=bw) +
    ggtitle(titlestr) + xlab(titlestr) + ylab("Density") + theme_classic() 
  p1
}

# plot the two conservations score prob sums
genemask = !(FullAnnotation$UCSC_RefGene_Name=="")
p_scores <- post_histogram(CpGscore_full$median,genemask,0.05,"CpG Conservation score") +
  theme(aspect.ratio=1/4,legend.position = "none") +
  scale_x_continuous(limits=c(-3,3),breaks=c(-3,-2,-1,0,1,2,3))
print(p_scores)
# ggsave("CpGscores_overlay.svg")

# plot gene scores and adjusted bootstrap distribution
p1 <- permed_data %>% ggplot(aes(x=score_median)) + geom_histogram(aes(y=..count..), binwidth=0.05, fill="lightgray",color = "grey50") + xlim(c(-1,2.5)) + theme_classic() + ylab("Number of genes") + theme(aspect.ratio = 1)
p2 <- permed_data %>% ggplot(aes(x=score_median_v3)) + geom_histogram(aes(y=..count..), binwidth=0.02, fill="lightgray", color="grey50") + xlim(c(0,1)) + theme_classic() + ylab("Number of genes") + theme(aspect.ratio = 1)
p_genes <- ggarrange(p1,p2,nrows=2,ncol=1)
# ggsave("gene_scores_bootstrap.svg")
```

\newpage
\subsection{Summary of neff and Rhat for each parameter}
```{r neffRhat, echo=FALSE, warning=FALSE}
# for each parameter (mu, betaT, sigmaP, sigmaPT, sigmaT, sigmaE, lp__)
# reporting: mean n_eff, min n_eff, # n_eff<500, %, # n_eff<100, %
n_eff_mat <-
  data.frame(
  mu = mu_full$n_eff,
  betaT = betaT_full$n_eff,
  sigmaP = sigmaP_full$n_eff,
  sigmaPT = sigmaPT_full$n_eff,
  sigmaT = sigmaT_full$n_eff,
  sigmaE = sigmaE_full$n_eff,
  lp = lp_full$n_eff
  ) %>% gather(param,value,mu:lp) %>% group_by(param) %>% summarize(
  mean = mean(value),
  min = min(value),
  s500 = sum(value < 500),
  p500 = 100 * mean(value < 500),
  s100 = sum(value < 100),
  p100 = 100 * mean(value < 100),
  .groups = "drop"
  ) %>% arrange(match(.$param,c("mu","betaT","sigmaP","sigmaPT","sigmaT","sigmaE","lp")))

kable(n_eff_mat, col.names=c("Parameter", "Mean n_eff", "Min n_eff", "# <500", "% <500", "# <100", "% <100"), digits = 2)

# again for Rhat
Rhat_mat <-
  data.frame(
  mu = mu_full$Rhat,
  betaT = betaT_full$Rhat,
  sigmaP = sigmaP_full$Rhat,
  sigmaPT = sigmaPT_full$Rhat,
  sigmaT = sigmaT_full$Rhat,
  sigmaE = sigmaE_full$Rhat,
  lp = lp_full$Rhat
  ) %>% gather(param,value,mu:lp) %>% group_by(param) %>% summarize(
  mean = mean(value),
  max = max(value),
  s11 = sum(value > 1.1),
  p11 = 100 * mean(value > 1.1),
  s101 = sum(value > 1.01),
  p101 = 100 * mean(value > 1.01),
  .groups = "drop"
  ) %>% arrange(match(.$param,c("mu","betaT","sigmaP","sigmaPT","sigmaT","sigmaE","lp")))

kable(Rhat_mat, col.names=c("Parameter", "Mean Rhat", "Max Rhat", "# >1.1", "% >1.1", "# >1.01", "% >1.01"), digits = 2)
cat("  \n")

# indices to filter out sites Rhat>1,1
QCinds <- lp_full$Rhat > 1.1
```

For the remainder of these results, a filter is applied to remove all sites with Rhat > 1.1 in lp (log-posterior likelihood of entire site fit). Another filter is later applied to remove all genes with <5 CpG sites.

\newpage
\subsection{Intra-gene conservation}
We examine intra gene conservation 2 ways: a table looking at gene regulatory regions, and a CpG-distance analysis

\subsection{Regulatory regions}
Table comparing different groups of CpG sites:
```{r Table1, echo=FALSE, warning=FALSE}

#, include=TRUE, results='asis', fig.keep='all', fig.show='asis', fig.height=3, fig.width=5

prepped_data %>%
  group_by(`Island relation` = Relation_to_UCSC_CpG_Island) %>%
  summarise_at(vars(sigmaP:sigmaT,score_median), mean, na.rm = TRUE) %>%
  dplyr::rename(ir = `Island relation`) %>%
  mutate(ir = case_when(is.na(ir) ~ "Sea",
                        str_detect(ir, "^S") ~ str_replace(ir, "^S_", "South "),
                        str_detect(ir, "^N") ~ str_replace(ir, "^N_", "North "),
                        TRUE ~ ir)) %>%
  dplyr::rename(`Island relation` = ir) %>%
  arrange(c(1,4,2,5,3,6)) %>% # for island, N shore, S shore, N shelf, S shelf, sea
  rename(`CpG Conservation Score`=score_median) %>%
  gt() %>%
  fmt_number(columns = vars(sigmaP, sigmaPT, sigmaT, `CpG Conservation Score`), decimals = 3) %>%
  tab_spanner(label = "Hierarchical Variances", columns = vars(sigmaP, sigmaPT, sigmaT)) %>% cols_align("center")

# stats for cpg island relations
# summarize mean of each group, do kruskal wallis and post hoc Tukey
my_data = prepped_data %>% select(score_median, Relation_to_UCSC_CpG_Island) %>%
  group_by(ir = Relation_to_UCSC_CpG_Island) %>%
  mutate(ir = case_when(is.na(ir) ~ "Sea",
                        str_detect(ir, "^S") ~ str_replace(ir, "^S_", ""),
                        str_detect(ir, "^N") ~ str_replace(ir, "^N_", ""),
                        TRUE ~ ir))
my_data %>%
  summarise_at(vars(score_median), mean, na.rm = TRUE)
kruskal.test(score_median ~ ir, data = my_data)
#install.packages("MultNonParam")
library(MultNonParam)
tukey.kruskal.test(my_data$score_median, my_data$ir, alpha = 0.001)

# stats for functional regions
summarize_conservation <- function(data,str) {
  indsin <- str_detect(data$UCSC_RefGene_Group, str) & !is.na(data$UCSC_RefGene_Group)
  indsout <- !str_detect(data$UCSC_RefGene_Group, str) | is.na(data$UCSC_RefGene_Group)
  if (str=="any") {
    indsin <- !is.na(data$UCSC_RefGene_Group)
    indsout <- is.na(data$UCSC_RefGene_Group)
  }
  data.frame(cons_in=mean(data$score_median[indsin], na.rm=TRUE),
         cons_out=mean(data$score_median[indsout],na.rm=TRUE),
         nsites=sum(indsin),
         pval=wilcox.test(data$score_median[indsin],data$score_median[indsout])$p.value)
}

bind_rows({prepped_data %>% summarize_conservation("any") %>% mutate(region="Gene-Associated")},
          {prepped_data %>% summarize_conservation("TSS1500") %>% mutate(region="TSS1500")},
          {prepped_data %>% summarize_conservation("TSS200") %>% mutate(region="TSS200")},
          {prepped_data %>% summarize_conservation("5'UTR|5UTR|5URT") %>% mutate(region="5'UTR")},
          {prepped_data %>% summarize_conservation("1stExon") %>% mutate(region="1stExon")},
          {prepped_data %>% summarize_conservation("Body") %>% mutate(region="Body")},
          {prepped_data %>% summarize_conservation("3'UTR|3UTR") %>% mutate(region="3'UTR")}) %>%
  select(c(region,cons_in,cons_out,nsites,pval)) %>%
  gt() %>%
  fmt_number(columns = vars(cons_in,cons_out,pval), decimals = 3)

# new stats
FRdata = data.frame(CpGscore = log2(prepped_data$sigmaP / prepped_data$sigmaT) ) %>%
  mutate(FR = pull(prepped_data,UCSC_RefGene_Group)) %>%
  mutate(FR=if_else(is.na(FR),"NA",FR)) %>%
  mutate(iTSS200 = str_detect(FR,"TSS200"),
         i5UTR = str_detect(FR,"5'UTR|5UTR|5URT"),
         iBody = str_detect(FR,"Body"),
         i3UTR = str_detect(FR,"3'UTR|3UTR"),
         iTSS1500 = str_detect(FR,"TSS1500"),
         i1stEx = str_detect(FR,"1stExon"),
         iExBnd = str_detect(FR,"ExonBnd"))
fr.lm = lm(formula = CpGscore ~ iTSS200 + i5UTR + iBody + i3UTR + iTSS1500 + i1stEx + iExBnd, data = FRdata)
#fr.lm = lm(formula = CpGscore ~ iTSS200 + i5UTR + iBody + i3UTR, data = FRdata)
summary(fr.lm)

library(rcompanion)
fr.srh = scheirerRayHare(formula = CpGscore ~ iTSS200 + i5UTR + iBody + i3UTR + iTSS1500 + i1stEx + iExBnd, data = FRdata)
```

\newpage
\subsection{Figure 1: CpG Distance Analysis}
Corrected for gene direction, based on Ensembl gene database \newline
```{r Fig1, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=3, comment=NA}
# load ensembl db
library(EnsDb.Hsapiens.v75) # ensembl annotation with gene position (hg19 version)
edb <- EnsDb.Hsapiens.v75
geneKeys <- keys(edb, keytype="GENENAME") %>% intersect(unique(gened_data$UCSC_RefGene_Name))
edbData<- AnnotationDbi::select(edb, keys=geneKeys, keytype="GENENAME",
                                columns=c("GENEBIOTYPE", "GENESEQSTART", "GENESEQEND", "SEQNAME", "SEQSTRAND")) %>%
  filter(SEQNAME %in% c(as.character(1:22),"X","Y")) %>%
  mutate(GENESTART = GENESEQSTART*(SEQSTRAND==1) + GENESEQEND*(SEQSTRAND==-1)) %>%
  select(-c(GENESEQSTART,GENESEQEND))

# calculate CpG distance based on Ensembl gene start position and direction
distanced_data <- gened_data %>%
  filter(!is.na(UCSC_RefGene_Name)) %>%
  filter(UCSC_RefGene_Name %in% geneKeys) %>%
  mutate(EDBCHR=edbData$SEQNAME[match(UCSC_RefGene_Name,edbData$GENENAME)],
         EDBSTART=edbData$GENESTART[match(UCSC_RefGene_Name,edbData$GENENAME)],
         EDBDIR=edbData$SEQSTRAND[match(UCSC_RefGene_Name,edbData$GENENAME)],
         EDBTYPE=edbData$GENEBIOTYPE[match(UCSC_RefGene_Name,edbData$GENENAME)]
         ) %>%
  filter(CHR==EDBCHR) %>%
  mutate(cpgdist = if_else(EDBDIR==1, MAPINFO-EDBSTART, EDBSTART-MAPINFO))

distance_length <- distanced_data %>%
  filter(!is.na(UCSC_RefGene_Name)) %>%
  select(CHR, UCSC_RefGene_Name, cpgdist, c(sigmaP:sigmaT,score_median:score_reg)) %>%
  mutate(location_bin = cut(cpgdist, c(seq(-1500, 20000, by = 100)), include.lowest = TRUE)) %>%
  group_by(location_bin) %>%
  summarise(mean_sigmaP = mean(sigmaP), mean_sigmaPT = mean(sigmaPT), mean_sigmaT = mean(sigmaT),  mean_score_median = mean(score_median), mean_score_median2 = mean(score_median2), n_obs = n(), .groups='drop') %>%
  mutate(location_bin = as.numeric(str_extract(location_bin, "(?<=,)(.*)(?=])")))

distance_length %>%
  drop_na() %>%
  pivot_longer(mean_sigmaP:mean_sigmaT) %>% 
  mutate(name = factor(name, c("mean_sigmaP","mean_sigmaPT","mean_sigmaT"), c("SigmaP","SigmaPT","SigmaT"))) %>%
  sample_frac() %>%
  ggplot(aes(location_bin, value, color = name, fill = name)) +
  geom_point(key_glyph = draw_key_rect) +
  ggforce::facet_zoom(location_bin < 5000) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(title = NULL,
       fill = NULL,
       color = NULL,
       x = "Distance from gene start site (binned to 100 bp)",
       y = "Average Sigma Value", 
       caption = "Figure 1: Average Sigma values of single CpGs as a function of position relative to gene.") +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2")
cat("\n")

# again for conservation scores
distance_length %>%
  drop_na() %>%
  pivot_longer(mean_score_median) %>% 
  mutate(name = factor(name, c("mean_score_median"), c("Score_median"))) %>%
  sample_frac() %>%
  ggplot(aes(location_bin, value, color = name, fill = name)) +
  geom_point(key_glyph = draw_key_rect) +
  ggforce::facet_zoom(location_bin < 5000) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(title = NULL,
       fill = NULL,
       color = NULL,
       x = "Distance from gene start site (binned to 100 bp)",
       y = "Average Conservation Score", 
       caption = "Figure 1: Average conservation scores of single CpGs as a function of position relative to gene.") +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2")
cat("\n")

# and look at number of CpG in each bin
distance_length %>%
  drop_na() %>%
  pivot_longer(n_obs) %>%
  mutate(name = factor(name, c("n_obs"), c("n_obs"))) %>%
  sample_frac() %>%
  ggplot(aes(location_bin, value, color = name, fill = name)) +
  geom_point(key_glyph = draw_key_rect) +
  ggforce::facet_zoom(location_bin < 5000) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(title = NULL,
       fill = NULL,
       color = NULL,
       x = "Distance from gene start site (binned to 100 bp)",
       y = "Number of CpGs", 
       caption = "Figure 1: Frequency of single CpGs as a function of position relative to gene start site.") +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2")
```

\newpage
\section{Gene Conservation Scores}
Genes are scored by average of all CpG sites within gene
\subsection{Example Genes: APC, TP53, TTN, HLA-A, GAPDH}
```{r examplegenes, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=2, comment=NA}
# for these genes, show the CpG site distribution, and add vertical lines for the CpG sites in the gene (gened_data?), then a bold yellow vertical line for the overall mean

plotGeneScore <- function(geneStr,norm) {
  CpGscores <- gened_data %>% filter(str_detect(UCSC_RefGene_Name,paste0("^",geneStr,"$"))) %>% pull(score_median)

  ggplot(prepped_data) + geom_histogram(aes(x=score_median, y=..count../nrow(prepped_data)), binwidth=0.05,fill="lightgray") +
    xlim(-3,3) + ggtitle(sprintf("%s: CpG scores (%i sites)",geneStr, length(CpGscores))) +
    xlab("CpG conservation score") + ylab("Freq.") + theme_classic() + theme(legend.position = "none") +
    #geom_vline(xintercept=CpGscores, color="red", alpha=0.5, size=0.2) +
    geom_histogram(data=data.frame(score=CpGscores),aes(x=score,y=norm*..count../length(CpGscores)),binwidth=0.05, fill="blue",alpha=0.5) +
    geom_vline(xintercept=mean(CpGscores), color="blue",linetype="dashed", size=1)
}

cat("\n")
#plotGeneScore("TP53",0.5)
plotGeneScore("TUBA1A",0.285)
#ggsave("genescore_TUBA1A.svg")
cat("\n")
plotGeneScore("TTN",0.37)
#ggsave("genescore_TTN.svg")
cat("\n")
plotGeneScore("HLA-A",0.26)
#ggsave("genescore_HLA-A.svg")

# stats for each gene
# do one sample wilcoxon rank sum
gened_data %>% filter(str_detect(UCSC_RefGene_Name,paste0("^TUBA1A$"))) %>% pull(score_median) %>% wilcox.test(mu = 0, alternative = "two.sided")
gened_data %>% filter(str_detect(UCSC_RefGene_Name,paste0("^TTN$"))) %>% pull(score_median) %>% wilcox.test(mu = 0, alternative = "two.sided")
gened_data %>% filter(str_detect(UCSC_RefGene_Name,paste0("^HLA-A$"))) %>% pull(score_median) %>% wilcox.test(mu = 0, alternative = "two.sided")
```

\subsection{COSMIC Cancer Genes}
```{r cancergenes, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=2, comment=NA}
# make gene-wise averages
gene_summarize <- function(data, id, value, fun = mean, ...) {
  grouped_data <- data %>% group_by({{id}})

  bind_cols(
    summarise_at(grouped_data, dplyr::vars({{value}}), fun, ...),
    count(grouped_data) %>% dplyr::ungroup() %>% dplyr::select(.data$n)
  )
}

gene_summary <- gened_data %>%
  filter(!is.na(UCSC_RefGene_Name)) %>%
  gene_summarize(UCSC_RefGene_Name, c(score_median))

cancergenes1 <- read_csv("/Users/kevinmurgas/Documents/Data+ project/Cancer gene lists/SangerCensus_COSMIC_CRC.csv")
genescores <- gene_summary %>% filter(UCSC_RefGene_Name %in% (cancergenes1 %>% pull(`Gene Symbol`))) %>% pull(score_median)
ggplot(gene_summary) + geom_histogram(aes(x=score_median, y=..count../nrow(gene_summary)), binwidth=0.05) + xlim(-3,3) + ggtitle(sprintf("Gene Scores of %i CRC Genes", length(genescores))) + xlab("Gene conservation score") + ylab("Freq.") + theme_classic() + theme(legend.position = "none") + geom_vline(xintercept=genescores, color="blue", alpha=0.5, size=0.2)
cat("\n")

# source 2
cancergenes2 <- read_csv("/Users/kevinmurgas/Documents/Data+ project/Cancer gene lists/CRCgenes_atlasgeneticsoncology.csv")
genescores <- gene_summary %>% filter(UCSC_RefGene_Name %in% (cancergenes2 %>% pull(CancerGenes))) %>% pull(score_median)
ggplot(gene_summary) + geom_histogram(aes(x=score_median, y=..count../nrow(gene_summary)), binwidth=0.05) + xlim(-3,3) + ggtitle(sprintf("Gene Scores of %i CRC Genes", length(genescores))) + xlab("Gene conservation score") + ylab("Freq.") + theme_classic() + theme(legend.position = "none") + geom_vline(xintercept=genescores, color="blue", alpha=0.1, size=0.2)

# compute % overlap of database genes and significant genes (By adj bootstrap)
n1 = sum(permed_data %>% filter(score_median_v3>0.95) %>% pull(UCSC_RefGene_Name) %in% (cancergenes1 %>% pull(`Gene Symbol`)))
n2 = sum(permed_data %>% filter(score_median_v3>0.95) %>% pull(UCSC_RefGene_Name) %in% (cancergenes2 %>% pull(CancerGenes)))

# plot violins of: all genes, COSMIC, AGO
perm_temp <- permed_data %>% select(c(UCSC_RefGene_Name, score_median,score_median_v3,n))
temp <- rbind(perm_temp %>% mutate(group="all"),
              perm_temp %>% filter(UCSC_RefGene_Name %in% (cancergenes1 %>% pull(`Gene Symbol`))) %>% mutate(group="COSMIC"),
              perm_temp %>% filter(UCSC_RefGene_Name %in% (cancergenes2 %>% pull(CancerGenes))) %>% mutate(group="AGO")
              )
temp %>% mutate(group = factor(group, levels = c("COSMIC", "AGO", "all"))) %>%
  ggplot() + geom_violin(aes(y=score_median,x=group,fill=group), draw_quantiles = 0.5) + scale_y_continuous(limits=c(-1,2),breaks = c(-1,0,1,2)) + theme_classic() + xlab("Database") + ylab("Gene Conservation Score")
# ggsave("database_violins.svg")

# try with histogram
plotDBScore <- function(geneStr,norm) {
  CpGscores <- gened_data %>% filter(str_detect(UCSC_RefGene_Name,paste0("^",geneStr,"$"))) %>% pull(score_median)

  ggplot(prepped_data) + geom_histogram(aes(x=score_median, y=..count../nrow(prepped_data)), binwidth=0.05,fill="lightgray") +
    xlim(-3,3) + ggtitle(sprintf("%s: CpG scores (%i sites)",geneStr, length(CpGscores))) +
    xlab("CpG conservation score") + ylab("Freq.") + theme_classic() + theme(legend.position = "none") +
    #geom_vline(xintercept=CpGscores, color="red", alpha=0.5, size=0.2) +
    geom_histogram(data=data.frame(score=CpGscores),aes(x=score,y=norm*..count../length(CpGscores)),binwidth=0.05, fill="blue",alpha=0.5) +
    geom_vline(xintercept=mean(CpGscores), color="blue",linetype="dashed", size=1)
}

```

\newpage
\section{Gene Set Conservation Analysis}
```{r load_gene_sets, include=FALSE}
# load in the C5 gene sets from MSigDB
library(GSA)
gmtfile <- file.path("/Users/kevinmurgas/Documents/Data+ project/EPIC data","c2.cp.v7.1.symbols.gmt")
genesets <- GSA.read.gmt(gmtfile)
```

```{r GSCA, echo=FALSE, comment=NA, fig.width=8}
# make a data.frame for gene sets and scores
nsets <- length(genesets[[1]])
GSEA_df <- data.frame(GeneSet=genesets[[2]][1:nsets], score=numeric(nsets),
                      num_gene = unlist(lapply(genesets[[1]][1:nsets],length)),
                      num_gene_cut = numeric(nsets))

# have to order all genes first by rank of conservation (conserved genes come first)
genenames = permed_data$UCSC_RefGene_Name
ngene = length(genenames)

# reorder by evidence rank
ind = order(-permed_data$score_median) # use ranked conservation score as evidence
evidence = seq(ngene,1,-1)
genenames = genenames[ind]

# plot evidence ranked: permed_data$score_median
ranks <- permed_data$score_median
names(ranks) <- permed_data$UCSC_RefGene_Name
barplot(sort(ranks, decreasing = T))

ptm <- proc.time()
for (iset in 1:nsets) {
  # take only genes that we have in full gene list
  set <- unlist(genesets[[1]][iset]) %>% intersect(genenames)
  GSEA_df$num_gene_cut[iset] = length(set)

  # skip if num_gene <10 or >500
  if (length(set)<10 | length(set)>500) {next}

  # compute running sum (as PDF first)
  runsum_in = numeric(ngene)
  runsum_out = numeric(ngene)
  for (i in 1:ngene) {
    if (genenames[i] %in% set) {
      # evidence = permed_data$median, or ranks
      runsum_in[i] = evidence[i]
      runsum_out[i] = 0
    } else {
      runsum_in[i] = 0
      runsum_out[i] = 1
    }
  }
  # convert to CDFs
  runsum_in <- cumsum(runsum_in)/sum(runsum_in)
  runsum_out <- cumsum(runsum_out)/sum(runsum_out)

  # compute K-S-like statistic
  v <- runsum_in - runsum_out
  score <- max(c(0,v)) + min(c(0,v))
  GSEA_df$score[iset] <- score
}
cat(proc.time()-ptm)
cat(sprintf("Processed %i genes, %i pathways. Top 50 paths:",ngene,nsets))

# insert significance bootstrapping

# make table of top scores
GSEA_df %>% arrange(-score) %>%
  select(score,num_gene,GeneSet) %>%
  slice(1:50) %>%
  mutate(GeneSet = gsub('(.{1,90})(\\s|$)', '\\1\n', GeneSet)) %>%
  gt()
```

```{r Exit, include=FALSE}
knit_exit()
```

``` {r enrich, include=FALSE}
# expression data
# downloaded file: E-MTAB-2836-query-results.tpms.tsv
# source: https://www.ebi.ac.uk/gxa/experiments/E-MTAB-2836/Results
expression_atlas <- read_tsv(here("data-raw", "E-MTAB-2836-query-results.tpms.tsv"), skip = 4, 
                             col_types = cols(
                               .default = col_double(),
                               `Gene ID` = col_character(),
                               `Gene Name` = col_character()
                             ))

# Gene Set Enrichment
tidy_gsea <- function(genes) {
  AnnotationDbi::mapIds(org.Hs.eg.db::org.Hs.eg.db, genes, 'ENTREZID', 'SYMBOL') %>%
    ReactomePA::enrichPathway()
}

enrich_sigmaP <- permed_data %>%
  #filter(sigmaP_v3 < quantile(sigmaP_v3, 0.05, na.rm = TRUE)) %>%
  filter(sigmaP_v3 < 0.05) %>%
  pull(UCSC_RefGene_Name) %>% tidy_gsea()
enrich_sigmaPT <- permed_data %>%
  #filter(sigmaPT_v3 < quantile(sigmaPT_v3, 0.05, na.rm = TRUE)) %>%
  filter(sigmaPT_v3 < 0.05) %>%
  pull(UCSC_RefGene_Name) %>% tidy_gsea()
enrich_sigmaT <- permed_data %>%
  #filter(sigmaT_v3 < quantile(sigmaT_v3, 0.05, na.rm = TRUE)) %>%
  filter(sigmaT_v3 < 0.05) %>%
  pull(UCSC_RefGene_Name) %>% tidy_gsea()
enrich_score_median <- permed_data %>%
  #filter(score_median_v3 > quantile(score_median_v3, 0.95, na.rm = TRUE)) %>%
  filter(score_median_v3 > 0.95) %>%
  pull(UCSC_RefGene_Name) %>% tidy_gsea()
```

\newpage
\subsection{Gene-average Parameter Distributions}
```{r Params_gene, echo=FALSE, fig.fullwidth=TRUE, warning=FALSE, comment=NA}
# make gene-wise averages
gene_summarize <- function(data, id, value, fun = mean, ...) {
  grouped_data <- data %>% group_by({{id}})

  bind_cols(
    summarise_at(grouped_data, dplyr::vars({{value}}), fun, ...),
    count(grouped_data) %>% dplyr::ungroup() %>% dplyr::select(.data$n)
  )
}

cat("Summarize by gene, taking mean value of all sites in gene")
ptm<-proc.time()
gene_summary <- gened_data %>%
  filter(!is.na(UCSC_RefGene_Name)) %>%
  gene_summarize(UCSC_RefGene_Name, c(mu:sigmaT,score_median))
cat(proc.time()-ptm)

post_histogram <- function(postvals,bw,xl,titlestr) {
  # stacked histogram colored by gene averages
  # and prior curve overlaid
  plot_df <- data.frame(val=postvals)
  p1 <- ggplot(plot_df, aes(x=val)) +
    geom_histogram(aes(y=..density..), binwidth=bw) +
    ggtitle(titlestr) + xlim(xl) + theme_classic() + xlab(titlestr) + ylab("Freq.")
  p1
}

# plot: n_sites, mu, betaT, sigmaP, sigmaPT, sigmaT
n_cutlo = 5
n_cuthi = quantile(gene_summary$n, 0.99)
pN <- post_histogram(gene_summary$n,1,c(0,200),paste0("N sites (max = ",max(gene_summary$n),")")) + geom_vline(xintercept=n_cutlo-0.5,linetype = "dashed",color="red") + annotate(geom="text",x=80,y=0.03,size=2,label=paste0("# genes <",n_cutlo," sites:\n", sum(gene_summary$n<n_cutlo)," / ",nrow(gene_summary))) +
geom_vline(xintercept=n_cuthi,linetype = "dashed",color="red") +annotate(geom="text",x=120,y=0.013,size=2,label=sprintf("# genes >%0.1f sites:\n %i / %i (99th pctile)", n_cuthi,sum(gene_summary$n>n_cuthi),nrow(gene_summary)))
gene_summary <- gene_summary %>% filter(n>=n_cutlo) %>% filter(n<n_cuthi)
pM <- post_histogram(gene_summary$mu,0.1,c(-5,5),"Mu")
pB <- post_histogram(gene_summary$betaT,0.1,c(-5,5),"BetaT")
pP <- post_histogram(gene_summary$sigmaP,0.05,c(0,2),"SigmaP")
pPT <- post_histogram(gene_summary$sigmaPT,0.05,c(0,2),"SigmaPT")
pT <- post_histogram(gene_summary$sigmaT,0.05,c(0,2),"SigmaT")
pP1 <- post_histogram(gene_summary$score_median,0.05,c(-3,3),"Score_median")
# arrange each parameter as a subplot
p_all <- ggarrange(pN,pM,pB,pP,pPT,pT,pP1,hjust=0,nrow=3,ncol=3)
print(p_all)
rm(pN,pM,pB,pP,pPT,pT,pP1)

```

\newpage
\subsection{Gene-average parameters vs N sites}
```{r Params_vsN, echo=FALSE, fig.fullwidth=TRUE, warning=FALSE, comment=NA, dev = "png"}
pM <- ggplot(gene_summary) + geom_point(aes(x=n,y=mu), alpha=0.01) + ggtitle("Mu vs N-sites")
pB <- ggplot(gene_summary) + geom_point(aes(x=n,y=betaT), alpha=0.01) + ggtitle("BetaT vs N-sites")
pE <- ggplot(gene_summary) + geom_point(aes(x=n,y=sigmaE), alpha=0.01) + ggtitle("SigmaE vs N-sites")
pP <- ggplot(gene_summary) + geom_point(aes(x=n,y=sigmaP), alpha=0.01) + ggtitle("SigmaP vs N-sites")
pPT <- ggplot(gene_summary) + geom_point(aes(x=n,y=sigmaPT), alpha=0.01) + ggtitle("SigmaPT vs N-sites")
pT <- ggplot(gene_summary) + geom_point(aes(x=n,y=sigmaT), alpha=0.01) + ggtitle("SigmaT vs N-sites")

pP1 <- ggplot(gene_summary) + geom_point(aes(x=n,y=score_median), alpha=0.01) + ggtitle("Score_median vs N-sites")

# arrange each parameter as a subplot
p_all <- ggarrange(pM,pB,pE,pP,pPT,pT,pP1,hjust=0,nrow=3,ncol=3)
print(p_all)
rm(pM,pB,pE,pP,pPT,pT,pP1)
```

\subsection{Figure 2: Bootstrapping}
```{r Fig2, echo=FALSE, warning=FALSE, fig.height=4, comment=NA}
# Figure 2
permed_data %>%
  pivot_longer(dplyr::matches("1|3")) %>%
  separate(name, into = c("type", "method"), sep = "_v") %>%
  mutate(method = factor(method, c("1", "3"), c("Naive bootstrap", "Adjusted bootstrap")),
         type = factor(type, c("sigmaP","sigmaPT","sigmaT"), c("SigmaP","SigmaPT","SigmaT"))) %>%
  drop_na() %>%
  ggplot(aes(value, fill = type)) +
  geom_histogram(bins = 50, color = "grey30") +
  facet_grid(method~type, scales = "free") +
  theme_minimal() +
  labs(x = "p-value",
       y = "Number of genes",
       title = NULL,
       caption = "Figure 2A: Distribution of boot-strapping p-values for genes as a function of sigma level and bootstrap method.")  +
  theme(legend.position = "none") +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2")

cat("***")
permed_data %>%
  pivot_longer(dplyr::matches("1|3")) %>%
  separate(name, into = c("type", "method"), sep = "_v") %>%
  mutate(method = factor(method, c("1", "3"), c("Naive bootstrap", "Adjusted bootstrap")),
         type = factor(type, c("score_median"), c("Score_median"))) %>%
  drop_na() %>%
  ggplot(aes(value, fill = type)) +
  geom_histogram(bins = 50, color = "grey30") +
  facet_grid(method~type, scales = "free") +
  theme_minimal() +
  labs(x = "p-value",
       y = "Number of genes",
       title = NULL,
       caption = "Figure 2B: Distribution of boot-strapping p-values for genes as a function of conservation score and bootstrap method.")  +
  theme(legend.position = "none") +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2")
```

\newpage
\subsection{Figure 3: Reactome Pathway Enrichment Analysis}
Enrichment defined from the group of genes <5th percentile of each sigma, or >95th percentile of conservation score.
```{r Fig3, echo=FALSE, fig.height=3.5, comment=NA}
cat("Number of enrichment pathways (sigmas):")
c(sigmaP = length(enrich_sigmaP$Description),
  sigmaPT = length(enrich_sigmaPT$Description),
  sigmaT = length(enrich_sigmaT$Description)) %>%
  sort()

list(
  sigmaP = enrich_sigmaP$Description,
  sigmaPT = enrich_sigmaPT$Description,
  sigmaT = enrich_sigmaT$Description
) %>%
  UpSetR::fromList() %>%
  UpSetR::upset(sets.bar.color = RColorBrewer::brewer.pal(3, "Set2")[c(3, 2, 1)])
grid::grid.text("Figure 3A: Frequency of overlap between pathways that are called \nas most conserved for each hierarchical variance.", x=0.65, y=0.05, gp=grid::gpar(fontsize=10))

cat("Number of enrichment pathways (probs):")
c(score_median = length(enrich_score_median$Description)) %>%
  sort()

list(
  score_median = enrich_score_median$Description
) %>%
  UpSetR::fromList() %>%
  UpSetR::upset(sets.bar.color = RColorBrewer::brewer.pal(3, "Set2")[c( 1)])
grid::grid.text("Figure 3B: Frequency of overlap between pathways that are called \nas most conserved for each conservation score.", x=0.65, y=0.05, gp=grid::gpar(fontsize=10))
```

\newpage
\subsection{Figure 4: Pathway Enrichment Network}
```{r Fig4, echo=FALSE, message=FALSE, fig.height=2.5}
# Figure 4
cat("sigmaP:")
ggg <- enrich_sigmaP %>%
  ReactomePA::emapplot(showCategory = 30, layout = "graphopt") +
  labs(caption = "Figure 4A: Relationship between pathways that are called as conserved based on sigmaP.", title = NULL)

ggg$layers <- ggg$layers[1:2]

ggg + coord_flip() + 
  ggraph::geom_node_label(aes_(label = ~name), repel = TRUE, size = 2) +
  scale_color_gradientn(colors = "black") +
  guides(color = "none", size = "none")

#ggsave(filename = here("figures", "figure-4.png"), width = 7, dpi = 320, height = 4.326)

# sigmaPT
cat("sigmaPT:")
ggg <- enrich_sigmaPT %>%
  ReactomePA::emapplot(showCategory = 30, layout = "graphopt") +
  labs(caption = "Figure 4B: Relationship between pathways that are called as conserved based on sigmaPT.", title = NULL)

ggg$layers <- ggg$layers[1:2]
ggg + coord_flip() + 
  ggraph::geom_node_label(aes_(label = ~name), repel = TRUE, size = 2) +
  scale_color_gradientn(colors = "black") +
  guides(color = "none", size = "none")

# sigmaT
cat("sigmaT:")
ggg <- enrich_sigmaT %>%
  ReactomePA::emapplot(showCategory = 30, layout = "graphopt") +
  labs(caption = "Figure 4C: Relationship between pathways that are called as conserved based on sigmaT.", title = NULL)

ggg$layers <- ggg$layers[1:2]
ggg + coord_flip() + 
  ggraph::geom_node_label(aes_(label = ~name), repel = TRUE, size = 2) +
  scale_color_gradientn(colors = "black") +
  guides(color = "none", size = "none")
```

\newpage
```{r Fig4DEF, echo=FALSE, message=FALSE, fig.height=2.5}
# score_median
cat("score_median")
ggg <- enrich_score_median %>%
  ReactomePA::emapplot(showCategory = 30, layout = "graphopt") +
  labs(caption = "Figure 4D: Relationship between pathways that are called as conserved based on score_median.", title = NULL)
ggg$layers <- ggg$layers[1:2]
ggg + coord_flip() + 
  ggraph::geom_node_label(aes_(label = ~name), repel = TRUE, size = 2) +
  scale_color_gradientn(colors = "black") +
  guides(color = "none", size = "none")

```

<!-- \newpage -->
<!-- \section{Patient-specific analysis} -->
<!-- Basic overview: using a specified number of genes (here 100), take top genes (here using score_prob bootstrapping percentiles) to select a "gene signature" of sorts. Then for each patient, calculate the mean pair-wise distance within the individual patients' tumor samples. Method 1 is these raw values, method 2 normalizes by the mean PWD of all normal samples. Method 3 just examines the overall mean PWD for all tumor samples, normalized by mean PWD of all normal. -->

<!-- ```{r PatientPWD, echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.height=9, fig.width=7} -->
<!-- library(gtools) -->

<!-- # pull out all samples beta values -->
<!-- all_beta <- FullAnnotation[,10:57] -->

<!-- # extract patient and tissue labels -->
<!-- patientLabel <- substr(colnames(all_beta), 1, 1) -->
<!-- patientLabel[10:12] <- "K*" -->
<!-- tissueLabel <- substr(colnames(all_beta), 2, 2) -->
<!-- tissueLabel[!(tissueLabel %in% "N")] <- "T"   #replace A/B/D/M with T -->

<!-- pats = patientLabel %>% sort %>% unique -->
<!-- npat = length(pats) -->

<!-- enrichedGenes <- permed_data %>% filter(score_prob_v3 < quantile(score_prob_v3, 0.05, na.rm = TRUE)) %>% -->
<!--   select(UCSC_RefGene_Name, score_prob, score_prob_v3, n) %>% -->
<!--   rename(Gene=UCSC_RefGene_Name, Pctile=score_prob_v3) -->

<!-- #ngene = 70 # using top N genes -->
<!-- #genes = enrichedGenes %>% top_n(-ngene,Pctile) %>% pull(Gene) -->
<!-- #genes = genes[1:ngene] -->

<!-- genes = enrichedGenes$Gene -->
<!-- ngene = length(genes) -->

<!-- pwd <- function(x) { -->
<!--   mean(dist(x)) -->
<!-- } -->
<!-- pwd2 <- function(x) { -->
<!--   abs(diff(x)) -->
<!-- } -->

<!-- signature_mat1 = matrix(,ngene,npat) # raw tumor pwd -->
<!-- signature_mat2 = matrix(,ngene,npat) # tumor pwd normalized by all normal pwd -->
<!-- for (igene in 1:ngene) { -->
<!--   # each gene, get gene inds, calculate PWD for the tumor samples -->
<!--   geneind = grep(genes[igene], FullAnnotation$UCSC_RefGene_Name) -->

<!--   # calculate PWD of each tumor -->
<!--   tum_pwd <- numeric(npat) -->
<!--   for (ipat in 1:npat) { -->
<!--     # each patient, get tumor sample inds -->
<!--     patind <- which(patientLabel == pats[ipat] & tissueLabel == "T") -->
<!--     tum_pwd[ipat] <- all_beta[geneind,patind] %>% logit %>% apply(1,pwd2) %>% mean -->
<!--   } -->

<!--   # calculate PWD of all normal -->
<!--   norm_pwd <- all_beta[geneind,tissueLabel == "N"] %>% logit %>% apply(1, pwd) %>% mean -->

<!--   signature_mat1[igene,] = tum_pwd -->
<!--   signature_mat2[igene,] = tum_pwd/norm_pwd -->
<!-- } -->

<!-- #cat("  \n\\newline  \nMethod 1 - raw intratumor PWD within each patient") -->
<!-- #signature_mat1 %>% as_tibble %>% `colnames<-`(pats) %>% gather %>% -->
<!-- #  mutate(gene=rep(genes,npat)) %>% rename(Patient=key,PWD=value) %>% -->
<!-- #  ggplot + geom_tile(aes(y=gene,x=Patient,fill=PWD)) + ylab(paste0(ngene," genes")) + -->
<!-- #  theme(axis.text.y = element_blank()) -->
<!-- #cat("  \n\\newline  \nMethod 2 - intratumor PWD normalized by all normal PWD") -->
<!-- #signature_mat2 %>% as_tibble %>% `colnames<-`(pats) %>% gather %>% -->
<!-- #  mutate(gene=rep(genes,npat)) %>% rename(Patient=key,PWD=value) %>% -->
<!-- #  ggplot + geom_tile(aes(y=gene,x=Patient,fill=PWD)) + ylab(paste0(ngene," genes")) + -->
<!-- #  theme(axis.text.y = element_blank()) -->
<!-- #cat("  \n\\newline  \nMethod 3 - all tumor PWD normalized by all normal PWD") -->
<!-- #signature_mat3 %>% as_tibble %>% mutate(gene=genes) %>% rename(PWD=V1) %>% -->
<!-- #  ggplot + geom_tile(aes(y=gene,x=1,fill=PWD)) + ylab(paste0(ngene," genes")) + -->
<!-- #  theme(axis.text.y = element_blank()) -->

<!-- library(pheatmap) -->
<!-- library(dendsort) -->

<!-- cat("  \n\\newline  \nMethod 1 - raw intratumor PWD within each patient") -->
<!-- df1 = as.data.frame(signature_mat1) -->
<!-- rownames(df1) <- genes -->
<!-- colnames(df1) <- pats -->
<!-- sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...))) -->
<!-- mat_cluster_cols <- sort_hclust(hclust(dist(t(df1)))) -->
<!-- mat_cluster_rows <- sort_hclust(hclust(dist(df1))) -->
<!-- pheatmap( -->
<!--   mat=df1, -->
<!--   cluster_cols = mat_cluster_cols, -->
<!--   cluster_rows = mat_cluster_rows, -->
<!--   show_rownames = FALSE, -->
<!--   show_colnames = TRUE, -->
<!--   main = "Top 1118 Genes: Tumor PWD" -->
<!-- ) -->

<!-- cat("  \n\\newline  \nMethod 2 - intratumor PWD normalized by all normal PWD") -->
<!-- df1 = as.data.frame(signature_mat2) -->
<!-- rownames(df1) <- genes -->
<!-- colnames(df1) <- pats -->
<!-- sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...))) -->
<!-- mat_cluster_cols <- sort_hclust(hclust(dist(t(df1)))) -->
<!-- mat_cluster_rows <- sort_hclust(hclust(dist(df1))) -->
<!-- pheatmap( -->
<!--   mat=df1, -->
<!--   cluster_cols = mat_cluster_cols, -->
<!--   cluster_rows = mat_cluster_rows, -->
<!--   show_rownames = FALSE, -->
<!--   show_colnames = TRUE, -->
<!--   main = "Top 1118 Genes: Tumor PWD" -->
<!-- ) -->
<!-- ``` -->

\newpage
\section{Appendix: Enrichment Gene/Pathway Lists}
\subsection{A1: SigmaP}
Genes are selected with cutoff <5th percentile (adjusted bootstrap), showing top 30 genes/pathways
```{r App_sigmaP, echo=FALSE, fig.fullwidth=TRUE, comment=NA, message=FALSE}
enrichedGenes <- permed_data %>%
  select(UCSC_RefGene_Name, sigmaP, sigmaP_v3, n) %>%
  rename(Gene=UCSC_RefGene_Name, Pctile=sigmaP_v3) %>%
  filter(Pctile < quantile(Pctile, 0.05, na.rm = TRUE)) %>%
  arrange(Pctile)

enrichedPaths <- tibble(Pathway = enrich_sigmaP$Description, ngenes = enrich_sigmaP$Count, padj = enrich_sigmaP$p.adjust) %>% arrange(padj)

ngene <- dim(enrichedGenes)[1]
npath <- length(enrich_sigmaP$Description)

cat(paste0("Total number of genes: ", ngene, "     Number of enrichment pathways: ", npath))

nshow = 30
if (ngene > nshow) {
  enrichedGenes <- enrichedGenes[1:nshow,]
}
if (npath > nshow) {
  enrichedPaths <- enrichedPaths[1:nshow,]
}

library(pander)
enrichedGenes %>% pander()

cat("SigmaP top 30 enriched paths:  \n")
enrichedPaths %>% pander()
```

\newpage
\subsection{A2: SigmaPT}
Genes are selected with cutoff <5th percentile (adjusted bootstrap), showing top 30 genes/pathways
```{r App_sigmaPT, echo=FALSE, comment=NA, message=FALSE}
enrichedGenes <- permed_data %>%
  select(UCSC_RefGene_Name, sigmaPT, sigmaPT_v3, n) %>%
  rename(Gene=UCSC_RefGene_Name, Pctile=sigmaPT_v3) %>%
  filter(Pctile < quantile(Pctile, 0.05, na.rm = TRUE)) %>%
  arrange(Pctile)

enrichedPaths <- tibble(Pathway = enrich_sigmaPT$Description, ngenes = enrich_sigmaPT$Count, padj = enrich_sigmaPT$p.adjust) %>% arrange(padj)

ngene <- dim(enrichedGenes)[1]
npath <- length(enrich_sigmaPT$Description)

cat(paste0("Total number of genes: ", ngene, "     Number of enrichment pathways: ", npath))

nshow = 30
if (ngene > nshow) {
  enrichedGenes <- enrichedGenes[1:nshow,]
}
if (npath > nshow) {
  enrichedPaths <- enrichedPaths[1:nshow,]
}

enrichedGenes %>% pander()

cat("\nSigmaPT top 30 enriched paths:  \n")
enrichedPaths %>% pander()
```

\newpage
\subsection{A3: SigmaT}
Genes are selected with cutoff <5th percentile (adjusted bootstrap), showing top 30 genes/pathways
```{r App_sigmaT, echo=FALSE, comment=NA, message=FALSE}
enrichedGenes <- permed_data %>%
  select(UCSC_RefGene_Name, sigmaT, sigmaT_v3, n) %>%
  rename(Gene=UCSC_RefGene_Name, Pctile=sigmaT_v3) %>%
  filter(Pctile < quantile(Pctile, 0.05, na.rm = TRUE)) %>%
  arrange(Pctile)

enrichedPaths <- tibble(Pathway = enrich_sigmaT$Description, ngenes = enrich_sigmaT$Count, padj = enrich_sigmaT$p.adjust) %>% arrange(padj)

ngene <- dim(enrichedGenes)[1]
npath <- length(enrich_sigmaT$Description)

cat(paste0("Total number of genes: ", ngene, "     Number of enrichment pathways: ", npath))

nshow = 30
if (ngene > nshow) {
  enrichedGenes <- enrichedGenes[1:nshow,]
}
if (npath > nshow) {
  enrichedPaths <- enrichedPaths[1:nshow,]
}

enrichedGenes %>% pander()

cat("SigmaT top 30 enriched paths:  \n")
enrichedPaths %>% pander()
```

\newpage
\subsection{A4: Score Median}
Genes are selected with cutoff >95th percentile (adjusted bootstrap), showing top 50 genes/pathways
```{r App_score_median, echo=FALSE, comment=NA, message=FALSE}
enrichedGenes <- permed_data %>% 
  select(UCSC_RefGene_Name, score_median, score_median_v3, n) %>%
  rename(Gene=UCSC_RefGene_Name, Score_median=score_median, Pctile=score_median_v3) %>%
  filter(Pctile > quantile(Pctile, 0.95, na.rm = TRUE)) %>%
  arrange(Pctile)

enrichedPaths <- tibble(Pathway = enrich_score_median$Description, ngenes = enrich_score_median$Count, padj = enrich_score_median$p.adjust) %>% arrange(padj)

ngene <- dim(enrichedGenes)[1]
npath <- length(enrich_score_median$Description)

cat(paste0("Total number of genes: ", ngene, "     Number of enrichment pathways: ", npath))

nshow = 50
if (ngene > nshow) {
  enrichedGenes <- enrichedGenes[1:nshow,]
}
if (npath > nshow) {
  enrichedPaths <- enrichedPaths[1:nshow,]
}

enrichedGenes %>% pander()

cat("Score_median top 50 enriched paths:  \n")
enrichedPaths %>% pander()
```